### 2016-03-17 copyright-page.tex
Add copyright page and ACH letterhead

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

% TODO: This is the only thing you need to change here each year!
\newcommand\thisyear{2016}

\multirow{3}{*}{\includegraphics[width=\textwidth]{copyright-header.png}}\\
\vspace{10em}

SIGBOVIK

A Record of the Proceedings of SIGBOVIK \thisyear

ISSN 2155-0166

April 1, \thisyear

\vspace{2em}

Copyright is maintained by the individual authors, though obviously this all gets posted to the Internet and stuff, because it's \thisyear.

Permission to make digital or hard copies of portions of this work for personal use is granted; permission to make digital or hard copies of portions of this work for classroom use is also granted, but seems ill-advised.
Abstracting with credit is permitted; abstracting with credit cards seems difficult.

Additional copies of this work may be ordered from Lulu; refer to {\tt http://sigbovik.org} for details.

\thispagestyle{empty}

\end{document}