### 2017-03-18 reviews/SIGBOVIK_2017_review-23.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Paper Review}} \\
&\\
& \LARGE{Paper 23: Distinguishing humans from} \\
& \LARGE{other forms of cattle} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

(MESSAGE FROM TRANSLATION COMMITTEE:
INCOMING REVIEW EXTREMELY UNORTHODOX IN COMPOSITION.
TRANSLATION INCLUDES MANY LINGUAL BEST-FITS.
FOR CLARITY, BEST-FITS ARE DENOTED BY ASTERISK PAIRS.
OVERALL ACCURACY OF REVIEW: UNKNOWN.
MESSAGE BEGINS...)

\newcommand\orz[1]{\textasteriskcentered{}#1\textasteriskcentered{}}

{\large\bf
\begin{tabular}{l}
\orz{Frumple}, \orz{Pretty Space Playground} \\
Knowledge feeling: High \\
Party feeling: \orz{Squeezing} the \orz{juice}! \\
\end{tabular}}
\vspace{1em}

Zoq-Fot-Pik are such silly \orz{campers}.
Human computer game YOLO is plain \orz{sister} of Androsynth \orz{learning machine}.
To begin, these words are must become \orz{singing} an appropriate \orz{song} in SIGBOVINE 2084.
Now you know the words.

But, this computer game is too \orz{lumpy} for silly Zoq-Fot-Pik.
Maybe you do not know.
Before the Ur-Quan \orz{dancing},
Humans and Androsynth \orz{advisors} together invent a \orz{learning machine} computer game,
but can not \orz{learn} it themselves.
Always after, computer games do not \orz{formal}, but have many \orz{enemies}.
Too many \orz{enemies} and the game will \orz{sliding}.
It is happiest days not to care about Androsynth anything.
Do not \orz{telling stories} about this computer game!

Any \orz{silly cow} sees these YOLO words to know person for Orz, but Orz is really \orz{fingers}! It is such a joke.
No more \orz{learning machine} is better.

\end{document}