### 2017-03-18 reviews/SIGBOVIK_2017_review-91.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Paper Review}} \\
&\\
& \LARGE{Paper 91: Fake News Logic} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Reviewer D\\
Ratings:  Absolutely great ratings, HUGE audience. \\
Confidence: I guarantee you there's no problem. Believe me.\\
\end{tabular}}
\vspace{1em}

The Philosophy Department's got great people.
They love me over there.
It's no CS department, but boy they've got a lot of computers.
My nephew would love it over there.
Will's a great guy, known him a long time, trustworthy, not like the media, believe me.

The global elite write a lot of papers on modal logic. They tell us how to think. Look what happened the last time we used one of those epistemic logics. Have you even seen their Kripke models? Their Kripke models remind of Rosie O'Donnell, 2/10, tops. Just sad. Your Kripke models are a real 10, just like my daughter. Great models.

I don't like all this talk about possible worlds, though.
We're getting killed at global trade, killed by Mexico, killed by China, killed by NATO.
Let's stop spending money on all these other possible worlds. Wasted money.
Let's spend that money on JOBS.

I know how to generalize this result to the syntactic setting.
I know a lot about syntax.
I've got the best syntax. Wharton grad.
Make $\mathcal{L}_{\mathit{FN}}$ \emph{strong}.
Lots of people are saying $S_5$ doesn't have a derivation system anyway.
They say we don't have a derivation system but then you go ask them ``Hey where's your derivation system'' and they don't have one!
Despicable.

\end{document}