### 2017-03-20 reviews/SIGBOVIK_2017_review-90.tex
Two new reviews

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Artifact Evaluation}} \\
&\\
& \LARGE{Paper 90: Automated Distributed Execution of} \\
& \LARGE{LLVM code using SQL JIT Compilation} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Anonymous first-year graduate student\\
Rating: Weak Reject \\
Confidence: Expert\\
\end{tabular}}
\vspace{1em}

I interpreted the abstract as a Piet\footnote{http://www.dangermouse.net/esoteric/piet.html}
program, downloaded the npiet interpreter\footnote{http://www.bertnase.de/npiet/}
and tried to run the abstract. I got the following error:

\begin{verbatim}
error: codelsize 597 does not match width of 600 pixel
\end{verbatim}

After cropping the image to be a 597x597 square (you should do this yourself!),
the program ran but produced the following trace:

\begin{verbatim}
info: verbose set to 1
info: trace set to 1
info: got 597 x 597 pixel with 256 cols
info: codelsize guessed is 597 pixel
trace: special case: we at a white codel - continuing

info: program end
\end{verbatim}

I'm not sure how to interpret this result.

\end{document}