### 2017-03-20 reviews/SIGBOVIK_2017_review-92.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\usepackage{microtype}
\setlength\parindent{0pt}
\setlength\parskip{12pt}
\usepackage{footnote}
\makesavenoteenv{tabular}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Paper Review}} \\
&\\
& \LARGE{Paper 92: Is it Percival time yet?:}\\
& \LARGE{A preliminary analysis of Avalon} \\
& \LARGE{gameplay and strategy}\\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
  Percival\\
  Rating: Vote your conscience\\
  Confidence: I'm so confused, guys\ldots
\end{tabular}}
\vspace{0.5em}

It's no fun when Merlin gets assassinated or evil wins, so Supposition
A sounds like something a spy would say. As a result, I question one
of this paper's fundamental premises. Is the author a spy?!
Regardless, this paper is a first-class analysis of our collective
obsession with Avalon.

{\large\bf
  \begin{tabular}{l}
    Percival II\\
    Vote: Approve\\
    Confidence: We've got to do this.
  \end{tabular}}
\vspace{0.5em}

I'm totally not evil, and I'm the \textit{real} Percival. I support
this mission.

{\large\bf
  \begin{tabular}{l}
    Ryan Kavanagh\\
    Rating: Approve\\
    Confidence: I think Percival is the real Percival?
  \end{tabular}}
\vspace{0.5em}

No more Percivals, please! Think of the poor stats sheet!

Normally we would recuse ourselves from reviewing papers we're
involved with,\footnote{Because SIGBOVIK is a serious conference with
  serious reviews.} but the program committee couldn't think of
anybody anybody who hadn't contributed to this paper's data set. This
paper provides a fun analysis of Avalon's innate funness. Section~4.5
claims there was insufficient data to perform a reliable chi-squared
analysis. Consequently, I encourage the author to switch her research
area to the analysis of Avalon and hire the program committee as
research assistants so that we can play more Avalon, err, I mean, help
her collect important data for a follow-up paper.

\end{document}