### 2017-03-21 reviews/SIGBOVIK_2017_review-14.tex
Two last-minute reviews

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Paper Review}} \\
&\\
& \LARGE{Paper 14: Is This the Shortest} \\
& \LARGE{SIGBOVIK Paper?}\\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Reviewer $\phi$\\
Rating: $e$/$\pi$\\
Confidence: $\epsilon$\\
\end{tabular}}
\vspace{1em}

While the importance of this work to the SIGBOVIK community can't be debated
(because we have no established protocol for such a debate), it is not novel.
The content of this paper is entirely contained within the papers
``The Portmantout'' from SIGBOVIK 2015,
``A shortmantout'' from SIGBOVIK 2016 and ``Efficient Computation of an
Optimal Portmantout'' from the present SIGBOVIK. It should therefore only
be accepted if our goal is to increase our page count which, as always, it is.

\end{document}