### 2017-03-21 reviews/SIGBOVIK_2017_review-79.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2017 Paper Review}} \\
&\\
& \LARGE{Paper 79: Dr.\ Boozehead, or How I learned to} \\
& \LARGE{stop worrying and get drunk: Design principles} \\
& \LARGE{and analysis of drinking games in the silicon age} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Robert J. Simmons, second general chair of the 2\textsuperscript{6}ish family of SIGsBOVIK \\
Rating: Unpleasantly sober \\
Confidence: Mostly shattered by JavaScript \\
\end{tabular}}
\vspace{1em}

There is yet again a drinking paper that doesn't cite my seminal work.
Kids these days.

\end{document}