### 2018-03-18 reviews/SIGBOVIK_2018_review_12.tex
reviews: Import corresponding to reviews-export-4

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper 12: Word\TeX} \\[0.25em]
& \LARGE{A WYSIPCTWOTCG Typesetting Tool} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Stephan Jenkins and Kevin Cadogan \\
Rating: Impressive but concerning \\
Confidence: 2/3 \\
\end{tabular}}
\vspace{1em}

The paper presents an impressive system for typesetting \TeX-style documents in Microsoft Word.
However, we are concerned with the validity of the double-blind study.
While reading the paper, we were able to see the labels on the table presenting the results.
A truly unbiased study would be triple-blind, hiding the labels not just from the subject and experimenter but also the reader.

\vspace{2em}
{\large\bf
\begin{tabular}{l}
Donald Knuth \\
Rating: Badness 10000 \\
Confidence: Overfull \\
\end{tabular}}
\vspace{1em}

I could just use Calibri and make my \TeX look like Word if I weren't typing this on a Mac.

\end{document}