### 2018-03-18 reviews/SIGBOVIK_2018_review_14.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper 14: Automating Art Snobbiness} \\[0.25em]
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{ll}
Simon Cowell\\
Rating: Thank you. No. \\
  Confidence: There is as much chance of this paper getting accepted \\
  as me flying to the moon tomorrow morning for breakfast. It's never \\
  going to happen.\\
\end{tabular}}
\vspace{1em}

If you had lived 2,000 years ago and done science like that, I think they would have stoned you.
The only decent thing about that paper was the end.

\end{document}