### 2018-03-18 reviews/SIGBOVIK_2018_review_24.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fdsymbol}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper 24: AAAAAAA} \\[0.25em]
& \LARGE{} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Returned, awaiting refund \\
Rating: $\medblackstar \medwhitestar \medwhitestar \medwhitestar \medwhitestar$ \\
Confidence: N \\
\end{tabular}}
\vspace{1em}

This product does not fit my device, and probably won't work with yours either.

AA was right out.  In fact, it was clear that my device needed something smaller than a AAA.
Sorting my Amazon results alphabetically and scrolling through more pages than I care to admit,
I was delighted to discover a product with this many As, and excited by the prospect of how small it
was bound to be!  Sure, it was unreviewed, but how could I have expected Amazon's highly relevant
search results to let me down?  Being somewhat of an online shopping connoisseur, I was unfazed when
I came home to discover a cardboard box larger than my washing machine and comprising 99.9\% packing
peanuts by volume eclipsing the entrance to my residence.  However, imagine my dismay when the
ensuing search-and-rescue operation revealed the contents to be the form factor of a few US letter
pages, far too big for my purposes.

Disgusted, I carefully replaced the unwanted item in its original packaging, placing that in turn
inside a refrigerator box for easier shipping, and mailed it back.  It's been three weeks, I'm still
awaiting my well-deserved refund, and I continue to suffer without being able to use my TV remote.

\end{document}