### 2018-03-18 reviews/SIGBOVIK_2018_review_EasierChair-2.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper EasierChair: A Survey of Hardw} \\[0.25em]
& \LARGE{are Multithreading} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Name Left Blank For Publication\\
Rating: \rotatebox{180}{$g$} \\
Confidence: Unwarranted\\
\end{tabular}}
\vspace{1em}

Did you know that Jacquard Looms were the first computers?
They ran on loops of punch cards, and punch cards and loops are things also associated with computers.
Surely any machine that can perform the same sequence of steps over and over -- like, say, a player piano or drinking bird -- must be Turing complete.
Some looms also had counters, and counting is the same as Turing completeness (didn't Alonzo Church show that?).

But let's get back to this great paper. Let me say that I really like it in theory, not that I'd ever do anything with fabric myself.
I mean, my inflexible age and gender identity -- along with a raft of biases -- prevent me from considering that something like this could be enjoyable or useful for people like myself.
Let me take a photo for my grandmother/wife/mother/girlfriend/daughter/New Mexico whiptail lizard; she'll love it.

I bought a pair of knit jeans the other day.
They were made with an extra-stretchy weave.

How soon will you add this feature to a loom so I can use it to knit a cut-and-sew shirt for my dog?
Can you make one from a 3D scan of my body? I just really want to send you a 3D scan of my body. Look, here's a dropbox link written on a napkin.

My t-shirt has a screen print of a houndstooth pattern on it.

Why aren't you doing something that really matters? Y'know, something the economy will reward you for, like extruding useless blobs of plastic, exploiting workers, or selling ephemeral digital tokens?

But seriously, nice work bro, definite publish.
See you at my ICO party later?

\end{document}