### 2018-03-18 reviews/SIGBOVIK_2018_review_17.tex
Import corresponding to reviews-export-5

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper 17: A Definitely Not Cherry-} \\[0.25em]
& \LARGE{Picked Rhetorical Analysis of} \\[0.25em]
& \LARGE{Programming Languages Reviews} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

\vspace{-2em}
{\large\bf
\begin{tabular}{l}
Reviewer A \\
Rating: Acce--uhhh, I mean, Weak Reject \\
Confidence: Untenured \\
\end{tabular}}
\vspace{-0.25em}

I liked this paper a lot.
It considers an important problem that---uhhh,
of course I think it could use more data,
yeah, no, I see what you're saying, yeah, I agree, let's reject this paper
I am on your team please write me a good letter.

{\large\bf
\begin{tabular}{l}
Reviewer B \\
Rating: Strong Accept \\
Confidence: \raisebox{0.075em}{\textless}3 \\
\end{tabular}}
\vspace{-0.25em}

I could not imagine a problem more important
to literally the entire scientific community.
This paper is beautifully written,
incorporating novel ideas into an impressive result.
Very nice!

{\large\bf
\begin{tabular}{l}
Reviewer C \\
Rating: Reject \\
Confidence: Expert \\
\end{tabular}}
\vspace{-0.25em}

Sorry to say, but I don't really see the point of this paper.
Why can't you just use the canonical mapping between letters and numbers?
The hard part of analyzing review rhetoric is
computing the means and standard deviations of statistics,
which this work fails to do.
Furthermore, the paper makes no actionable reccomendations for authors,
who wish to understand the review process in order to get papers accepted,
not merely to understand the rhetorical habits of reviewers.

{\large\bf
\begin{tabular}{l}
Reviewer D \\
Rating: Strong Reject \\
Confidence: $\infty$ \\
\end{tabular}}
\vspace{-0.25em}

Footnote citations, strong reject.

\end{document}