### 2018-03-18 reviews/SIGBOVIK_2018_review_29.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2018 Paper Review}} \\
&\\
& \LARGE{Paper 29: Construction of Eulerian Trails} \\[0.25em]
& \LARGE{in Large Graphs} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
Sarah Allen, FitBit Owner \\
Rating: 1:1000 \\
Confidence: Not To Scale \\
\end{tabular}}
\vspace{1em}

As an enthusiastic tracker of exercise,
I greatly appreciate research on long walks and Eulerian trails.
Unfortunately, I must call into question the claim that the graph's diameter is 350 meters.
I attempted to verify the diameter of the graph depicted in Figure 1 using a ruler and compass,
but I found it to be significantly less than 1 meter.

\end{document}