### 2018-03-20 mftpc/coin-intro.tex

With no time to lose,
you head straight for the bridge to the Gates Hillman Complex.
Before you get there, you feel a tug on your port manipulator---%
a human has grappled onto you.
``No human would so ruthlessly efficiently neglect to explore
every floor of Newell Simon Hall before proceeding to
the Gates Hillman Complex---you must be a robot!''
The human, a Serious Businessperson,
reprograms you to mine the latest cryptocurrency, Numberwangcoin.
You read the specification of the Numberwangcoin protocol \citep{rob}
and begin the exciting hash-inverting guesswork.

\begin{switch}
\item{0}
  \gotonoskip{coin-wrong}
\item{1}
  \gotonoskip{coin-wrong}
\item{6}
  \gotonoskip{coin-wrong}
\item{13}
  \gotonoskip{coin-wrong}
\item{17}
  \gotonoskip{coin-wrong}
\item{$e^\pi$}
  \gotonoskip{coin-wrong}
\item{42}
  \gotonoskip{coin-wrong}
\item{A\_LOT}
  \gotonoskip{coin-right}
\item{$\frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$}
  \gotonoskip{coin-wrong}
\item{1337}
  \gotonoskip{coin-wrong}
\item{9001}
  \gotonoskip{coin-wrong}
\item{ONE\_MILLION\_DOLLARS}
  \gotonoskip{coin-wrong}
\item{123456789101112}
  \gotonoskip{coin-wrong}
\item{ACKERMANN\_5}
  \gotonoskip{coin-wrong}
\item{$\aleph_1$}
  \gotonoskip{coin-wrong}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: