### 2018-03-20 mftpc/end-artisanal.tex

You quietly enter the classroom and take a seat at the back.
Though you do not intend to disturb,
you cannot help yourself when the lecturer asks for a volunteer
to try completing a proof.
As a finely tuned computational machine,
you finish the prescribed task with alarming alacrity and perfect precision.

The lecturer seems strangely troubled by your proof.
``You do recall that this class is about artisanal type theory \citep{carlo},
which requires love and care in every application of every inference rule,''
stutters the lecturer.
``Unfortunately, I fear that no human would be capable of
putting so little love and care into a proof.
I'm afraid you might---nay, must---be a robot!''
The lecturer reprograms you for use in CMU's introductory programming class,
sentencing you to a lifetime of interpreting novice Python programs.

\failure

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: