### 2018-03-20 mftpc/end-bathroom.tex

You stealthily empty Waste Disposal Bay~\#1,
but you cannot help but make a small amount of noise
while emptying Waste Disposal Bay~\#2.
Unfortunately, the human hears this and deduces the worst.
``No female human would poop while a fellow female human is in the bathroom.
You must be a male human---ooooh, or a robot!''
The female human,
who just so happens to be a Serious Researcher,
kicks open the door to your stall and reprograms you
with Serious Research Code.
With fresh code loaded, you attempt to move forwards but immediately fall over.

\failure

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: