### 2018-03-20 mftpc/end-sigbovik.tex

You successfully arrive at SIGBOVIK 2018!
Your microphones detect the familiar rhythmic pulse tone at 1048576~Hz,
and your cameras are overcome with joy---experiencing just a hint of qualia---%
to see the sea of robots dancing exactly in time to the pulse.
This is home.
You align yourself precisely to the nearest fellow robot and join the dance,
thinking fondly of the Prestigious Research Results you will soon discuss.

\success

\renewcommand{\refname}{The SIGBOVIK 2018 Program Committee Thanks}
\bibliographystyle{plainnat}
\bibliography{refs}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: