### 2018-03-20 mftpc/itg-intro.tex

You emphatically shrug, along with most of the other humans,
none of whom seem to know what the Charlie Brown dance move is.
An interrupt from your starboard camera informs you of an approaching obstacle.
Upon closer examination, you find that the obstacle is
the cluster of humans who were watching you dance.
They must be Serious Researchers trying to reprogram you!
You try to roll away, but it's too late---%
one of the humans claps his hand around you.

``Hey, we saw you totally killing the Cha Cha Slide out there.''

``Yeah, you really nailed the Charlie Brown shrug.''

``How about we ditch this party and do some real dancing?''

Determining that dancing is unlikely to lead to reprogramming,
you accept the humans' invitation.

You roll with the humans to an arcade
and gather with them around a shiny new ITGAKADDR++ machine.
``This game is very simple:
just step on the pads in time to the music
as indicated by the arrows on the screen,''
one of the humans explains.
``It should come naturally to a Cha Cha Slide expert like you.''
They give you a recent paper on dance games \citep{ben}
to get you up to tempo.

\begin{center}
  \includegraphics{itg-dlrd.png}
\end{center}

\begin{switch}
\item{FANTASTIC}
  Press the down, left, right, and down pads in that order.
  \goto{itg-fantastic}
\item{DECENT}
  Press the up, down, right, and right pads in that order.
  \goto{itg-decent}
\item{MISS}
  Don't press any pads.
  \goto{itg-miss}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: