### 2018-03-20 mftpc/pants-multi.tex

You search the room in vain for thread-like objects.
Eventually, you realize that the thread-like objects were inside you all along:
wires!
While unscrewing your main casing,
you load and analyze your wiring diagram.
You look down upon your own wires,
identify one that your analysis determined is not critical to your functionality,
carefully place it into the jaws of the scissors from your sewing kit,
and cu---

\segfault

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: