### 2018-03-20 mftpc/pants-single.tex

With mechanical precision, you easily repair your pants using a single thread.
Multithreaded pant repair must only be necessary for humans,
who could never stitch so precisely.

As you put your pants back on,
you notice that you are surrounded by curious observers.
Some look shocked that you so cavalierly removed your pants,
while others look impressed at your pant repair prowess.
Just in case any of the onlookers are Serious Researchers,
you scurry outside before any of them realize you are a robot.

\mustgoto{trail-intro}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: