### 2018-03-20 mftpc/slide-dance.tex

You rotate your wheels a quarter turn and boogie backwards,
along with the rest of the humans.
The following instructions,
starting with ``one hop this time'', are much more straightforward.
Seeing that the other humans are having so-called fun,
you set your top-mounted LCD to \texttt{:)} in solidarity.
On your starboard camera, you notice a small cluster of humans
looking at you and yelling furtively at each other,
but your microphones can't pick up their communication over the music.
You hope they are not Serious Researchers who suspect you of being
a robot ripe for reprogramming.

This thread is interrupted by an interrupt
from your natural language coprocessor,
which has reported zero possible meanings for the current instruction,
``Charlie Brown''.

\begin{switch}
\item{KICK\_THE\_FOOTBALL}
  Download a video on how to do the Charlie Brown dance move,
  watch it at 200-times speed,
  and attempt to execute the move.
  \goto{pants-intro}
\item{GOOD\_GRIEF}
  Hope that humans also have no idea what the Charlie Brown dance move is
  and emphatically shrug.
  \goto{itg-intro}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: