### 2018-03-20 mftpc/trail-elevator.tex

You press the 4~button, and the elevator starts moving.
The elevator makes a stop on floor~3
for the few thousand people taking the elevator up one floor to class.
``While we wait,
perhaps you can help me with a mathematics problem
that has been floating in the back of my head for a bit,''
says CoBot.
``What is $0.1 + 0.2$?''
This is not your first rodeo:
as a robot with an FPU in place of an ALU,
you are well aware of the perils of rounding errors.
Fortunately, a recent paper describes
a library for performing integer operations
using only floating-point assembly instructions \citep{jim_tom},
and you had the library installed in your latest upgrade.

\begin{switch}
\item{INT\_COMES\_NATURALLY}
  Use your new integer operations to reason
  $0.1 + 0.2 = (1 + 2) * 0.1 = 0.3$.
  \goto{trail-int}
\item{IEEE\_754EVER}
  Use your classic FPU to report
  $0.1 + 0.2 = 0.30000000000000004$.
  \goto{trail-float}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: