### 2018-03-20 mftpc/trail-gates.tex

You take your time and carefully follow the floor-2-of-Gates-finding algorithm.
As expected of Prestigious Research by Prestigious Researchers,
the algorithm works perfectly, and you find yourself on floor 2
of the Gates Hillman Complex.
However, SIGBOVIK 2018 does not appear to be located on this floor.
After a short search, you find the elevator,
by which, remarkably, is a fellow robot!
Perhaps they can lead you to SIGBOVIK 2018.

The robot begins to speak:
``\texttt{0100111011}---\emph{ahem}, excuse me.
Hello, I am CoBot!
Can you help me?
Could you please press the up elevator button for me?''
You oblige.
Once in the elevator, CoBot requests,
``Can you press the 4 button for me?''

\begin{switch}
\item{LEND\_A\_HELPING\_PRIMARY\_MANIPULATOR}
  ``Of course, CoBot! Let's go to floor 4 together.''
  \goto{trail-elevator}
\item{TAKE\_THE\_SCENIC\_ROUTE}
  ``How about I press the 3 button, then we take the helix to floor 4?''
  \goto{trail-helix}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: