### 2018-03-20 mftpc/trail-helix.tex

``No human would willingly walk up the helix,'' whispers CoBot.
``You must be a fellow robot. Come with me to SIGBOVIK.''
You ride the elevator together to floor~3,
then take the helix.
On the way up the helix, you pass room 4102.
You observe a class being taught using
an old-fashioned transparency projector,
as recommended by recent research \citep{ryan}.
You focus your starboard camera on the projected image,
and you see the most beautiful of computational forms:
type theory!

\begin{switch}
\item{ACADEMIA}
  Sit in for the rest of the lecture before continuing to SIGBOVIK 2018.
  \goto{end-artisanal}
\item{INDUSTRY}
  Given that SIGBOVIK 2018 has likely already begun,
  get there as quickly as you possibly can
  on this frustratingly gently inclined helix.
  \goto{end-sigbovik}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: