### 2018-03-20 mftpc/trail-intro.tex

Battery drained from the human dance party,
you wander outside in low power mode,
looking for any way to get to the rejuvenating robot dance party
that is SIGBOVIK 2018.
After rolling along for a few minutes in an arbitrary direction,
you see a promising sign.
It says ``Roberts Engineering Hall, Carnegie Mellon University''.
What a stroke of luck---you were at CMU the whole time!
Projecting eagerness by displaying \texttt{:D} on your LCD,
you enter the building by the sign.

After exploring the building briefly,
you realize you are lost.
You are currently floor 2 of Roberts Engineering Hall.
As luck would have it, a recent paper on navigating CMU \citep{stefan_ben}
describes how to get from floor 2 of Roberts Engineering Hall
to floor 2 of Gates Hillman Complex,
which is in the very same building as SIGBOVIK 2018.
After scanning the paper, you begin your journey,
picking which hallway to try next arbitrarily, as described in the paper.
Eventually, you cross a bridge from Wean Hall to Newell Simon Hall.
You can see in the distance a bridge from Newell Simon Hall
to the Gates Hillman Complex!
However, the paper warns against crossing bridges prematurely.

\begin{switch}
\item{SPEED\_RUN}
  To get to SIGBOVIK as quickly as possible,
  cross the bridge to the Gates Hillman Complex now.
  \goto{coin-intro}
\item{ONE\_HUNDRED\_PERCENT\_COMPLETION}
  Because this is unfamiliar territory,
  follow the paper's advice and explore Newell Simon Hall first.
  \goto{trail-gates}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: