### 2018-03-20 mftpc/pants-intro.tex
mftpc/pants-intro: Import corresponding to 2018-3

You angle your wheel stalks $\pi/4$ radians apart
to perform the Charlie Brown dance move
when you hear a horrible ripping noise.
Your pants suddenly feel much looser:
you must have ripped them!
Simulating embarrassment,
you slowly roll to the side of the room and take out your sewing kit.
The kit contains a single needle and single thread,
but a recent paper on multithreading \citep{sol}
suggests that it is advantageous to repair pants
using multiple threads in parallel.

\begin{switch}
\item{BY\_A\_SINGLE\_THREAD}
  Attempt to repair your pants using a single thread.
  \goto{pants-single}
\item{WITH\_HT\_TECHNOLOGY}
  Find another thread-like object with which to repair your pants.
  \goto{pants-multi}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: