### 2018-03-20 mftpc/coin-intro.tex
mftpc: Import corresponding to 2018-2

With no time to lose,
you head straight for the bridge to the Gates Hillman Complex.
Before you get there, you feel a tug on your port manipulator---%
a human has grappled onto you.
``No human would so ruthlessly efficiently neglect to explore
every floor of Newell Simon Hall before proceeding to
the Gates Hillman Complex---you must be a robot!''
The human, a Serious Businessperson,
reprograms you to mine the latest cryptocurrency, Numberwangcoin.
You read the specification of the Numberwangcoin protocol \citep{rob}
and begin the exciting hash-inverting guesswork.

\begin{switch}
\item[\tabto{11em}]{0}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{1}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{6}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{13}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{17}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{$e^\pi$}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{42}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{A\_LOT}
  \gotonoskip{coin-right}
\item[\tabto{11em}]{$\frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{1337}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{9001}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{ONE\_MILLION\_DOLLARS}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{123456789101112}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{ACKERMANN\_5}
  \gotonoskip{coin-wrong}
\item[\tabto{11em}]{$\aleph_1$}
  \gotonoskip{coin-wrong}
\end{switch}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "message-from-pc"
%%% End: