### 2019-03-17 reviews/SIGBOVIK_2019_review_1.tex
Import corresponding to reviews-2019 tag

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2019 Paper Review}} \\
&\\
& \LARGE{Paper 1: Monetization of Development Tools} \\[0.25em]
& \LARGE{for Fun and Profit} \\
\\
\hline
\end{tabular}}
\vspace{1em}
\thispagestyle{empty}

{\large\bf
\begin{tabular}{l}
ReviewIt Free\\
Rating: [Wait remaining: 2h35m; Reveal Now: \underline{25R\$}]\\
Confidence: [Upgrade Your Review Experience With Confidence Levels: \underline{75R\$}]\\
\end{tabular}}
\vspace{1em}

This paper has been reviewed using ReviewIt Free, a service of ReviewIt enterprises.
ReviewIt Free provides the same great review feedback to authors as ReviewIt Enterprise, ReviewIt Premium, ReviewIt Pro, ReviewIt Gold, and ReviewIt Solid Gold with a few limitations:
\vspace{-1em}
\begin{itemize}
\item first, all reviews begin with this block of text and a short but mandatory advertisement;
\item second, authors that wish to immediately view a numerical summary score of their review can do so for a small micro-payment -- the score will otherwise be revealed several hours after the review release, but before the committee needs it to make accept/reject decisions;
\item third, the reviewer confidence rating is also available for a larger micropayment.
\end{itemize}

These limitations ensure that ReviewIt enterprises can continue to provide great reviews to all our customers.
We have carefully considered all of our micropayments and decided that any diffuse negative influence they may have on the progress of science is certainly more than made up for by the concentrated positive influence they have on the balance in our wallets.\\

\hrule{}
\begin{center}
\framebox{\begin{minipage}{4in}
\textbf{Ad-Blocker Detected! No review text displayed.}
Please disable your ad-blocker so that we can force you to watch a five minute edutisement from PraegerU;
monthly subscribers don't need to watch these advertisements.
\end{minipage}
}
\end{center}

\hrule{}

Hey authors, not happy with your review results?
Upgrade your reviewing experience with these extra features:
\vspace{-1em}
\begin{itemize}
\item Proper grammar, spelling, and complete sentences. [\underline{100R\$}]
\item No statements of opinion couched as fact. [\underline{200R\$}]
\item Sensible feedback that considers that paper authors are people, too, and are generally making a good-faith effort to move science forward. [\underline{currently unavailable}]
\item Specific, constructive feedback. [\underline{10000R\$}]
\item Re-roll the review hoping for better drops. [\underline{submit the paper next year}]
\end{itemize}

\end{document}