### 2019-03-19 reviews/SIGBOVIK_2019_review_19.tex
Pull in Stefan's review of paper 19

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}
\usepackage{xcolor}
\usepackage{listings}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \Large{\em CONFIDENTIAL COMMITTEE MATERIALS} \\
&\\
& \textbf{\Huge{SIGBOVIK 2019 Paper Review}} \\
&\\
& \LARGE{Paper 19: Applications of} \\[0.25em]
& \LARGE{Standard ML at Google} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\pagestyle{empty}

{\large\bf
\begin{tabular}{l}
  Stefan Muller\\
  Former 15-150 Instructor\\
  Expensive Building, Forbes Avenue\\
Rating: 42.0 \\
Confidence: \texttt{val it = 42.0 : real}\\
\end{tabular}}
\vspace{1em}

This paper gives an excellent summary of the many benefits of Standard ML,
a strongly-typed, usable, production-quality language, as well as a
number\footnote{zero is a number} of applications of it in practice in
Google, a reasonably extensive search engine codebase developed by researchers
at Stanford.

I agree with the author that the publication of this study could be a great
learning experience to CS students at Carnegie Mellon, as well as a number
of other universities that make extensive use of Standard
ML\footnote{zero is a number}.

As future work, the author may wish to consider even further expanding the
use of Standard ML at Google. Its strong abstraction properties, for example,
could be useful to help Google avoid the privacy issues that have plagued
other tech companies in recent days. Consider the code on the following page.

\lstset{  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
  numbers=left
}

\clearpage
\begin{lstlisting}[language=ML]
signature User =
sig
    val name           : string
    val age            : int
    val interests      : string list
    val docs           : string list
    val sheets         : string Array.array Array.array list
    val search_history : string list
    val emails         : string list
    val contacts       : (string * int) list
end

signature AdUser =
sig
    (* All the info I want Google to give to advertisers *)
end

structure Me :> AdUser =
struct
...
end

functor Ads (U : AdUser) =
struct
...
end
\end{lstlisting}

\textit{Conflict of interest disclosure}: The reviewer, while generally
distrustful of Big Tech, uses Google products when necessary (read: always).

\end{document}