### 2020-03-08 reviews/SIGBOVIK_2020_review_5.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Reviewer Number Two-and-a-half}
\rating{Humbug, all of it!}
\confidence{At present, all things considered -- and especially my vast back-\\ground in computational heresy -- I must say that I am about as confident\\ regarding this review, as my uncle is that the quotes he posts on Facebook\\ are indeed Einstein's. In other words, I am absolutely bloody$^*$ confident.}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

After carefully not-reading the paper, I have convinced myself that the experiments are probably lacking. Honestly, which paper would not merit from more extensive empirical verification that nonetheless fits in the same amount of space? This is just a well-known fact about the academic space-time continuum, and can therefore only be true for this paper as well.

Another key issue with the paper is that it does not seem to cite \emph{any} of my own or my research group's papers! And I must ask: how on earth could the authors possibly have ignored such a key piece of related work? Our papers should be cited at least three times in any paper on this topic.

Finally, I would like to point out that the paper is ridden with spelling mistakes. Obviously, I am not asking the authors to be perfect, but ending perfectly innocent verbs with ``-ize'' instead of ``-ise'' and nouns with ``-ization'' instead of ``-isation''? Depravity!

I cannot do anything but conclude that this paper is complete and utter humbug, and should be read exclusively by those who wish to have a good laugh.

\vfill
$^*${\small Yes, ``bloody'', partially because I am indeed from across the pond, but mainly because I just slaughtered your paper. No, I am not in the least bit apologising; in fact I hereby offer the authors my two-fingered salute. Cheerio!}

\end{document}