### 2020-03-08 reviews/SIGBOVIK_2020_review_6.tex
Jenny's review exploit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}
\usepackage{lastpage}
\usepackage{secure-page-numbers}
\newcommand{\RNum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Anonymous}
\rating{lukewarm, getting warmer}
\confidence{Confident that the organizing committee are FOOLS}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
The paper is acceptable, if nearsighted in its understanding of the implications of its findings. HOWEVER the authors are NO WHERE as nearsighted as the organizing committee, which has provided me with the perfect vector for injecting as many page numbers as I please into the proceedings, allowing me to harass the proceedings chair under the full cover of anonymity. Behold, as I inform \emph{everyone} that this is Page \thepage \hspace{1pt} of \pageref{LastPage}, equivalent to \romannumeral \thepage\ in roman numeral, or \shalookup \thepage\ under the numbering scheme described in Cryptographically Secure Page Numbering in \LaTeX! Haha! Ha! Muahahaha! \end{document}