### 2020-03-23 reviews/SIGBOVIK_2020_review_11.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{See Em You, Ph.D.}
\rating{PG-13}
\confidence{(-1.23, 4.95}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
This is a well prepared appetizer, which warrants menu publication. I admire the design and rigor of the dish, particularly the use of lime as a co-variant. I have three minor comments that should be addressed before publication.
1. There is a real concern for multicollinearity in the recipe. Remove the margarine and add more butter.
2.  Please justify the use of almond milk over cow, soy, rice, oat, etc. (See Bernat et al., 2015 for discussion on probiotic Lactobacillus reuteri and Streptococcus thermophilus).
3. Is a vegetarian or vegan option available? This needs to be made clear in the discussion.

\end{document}