### 2020-03-23 reviews/SIGBOVIK_2020_review_9.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Reviewer 5/7}
\rating{5/7}
\confidence{Supercharged Chevy Silverado}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
%Only super serious reviews here.
%You don't want to impede the scientific process, do you?
We believe the subject of this paper is of utmost importance to the field, and appreciate its timeliness, timeline, timbre, and temerity. 
That being said, in order to be accepted into this prestigious venue, the we suggest the following changes:
\begin{itemize}
    \item The main methodology is seriously flawed and must be addressed.
    \item The authors used the word ``whilst'' 11 times. We recommend a minimum ``whilst'' usage count of 20 to be seriously considered for publication.
    \item The English language writing in this paper contains numerous spelling and grammar errors. Consider kidnapping a native English speaker, chaining them in your basement, and depriving them of food until the draft is written in language that William Strunk would be proud of.
    \item Please use raster rather than vector images for all figures, so that the reader may count the pixels used in the each letter of the text in the axis labels.
    \item In order to give due credit to all contributors to this work, please make Reviewer 5/7 a co-author.
    \item On page 44, line 13, eight word, second character, we found the letter ``o.'' We like this letter, please leave it as is.
    \item Additional changes have been suggested in the attached PDF.

\end{itemize}

\end{document}