### 2020-03-30 reviews/SIGBOVIK_2020_review_3.tex
reviews: Format Google ToS review properly

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{\textless sigbovik@gmail.com\textgreater}
\rating{\textless noreply-utos@google.com\textgreater}
\confidence{Learn more about our updated Terms of Service}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
We’re improving our Terms of Service and making them easier for you to
understand. The changes will take effect on March 31, 2020, and they won’t
impact the way you use Google services.

For more details, we’ve provided a summary of the key changes and Frequently
Asked Questions. At a glance, here’s what this update means for you:

\begin{itemize}
\item \textbf{Improved readability:} While our Terms remain a legal document, we’ve
done our best to make them easier to understand, including by adding
links to useful information and providing definitions.

\item \textbf{Better communication:} We’ve clearly explained when we’ll make
changes to our services (like adding or removing a feature) and when
we’ll restrict or end a user’s access. And we’ll do more to notify you
when a change negatively impacts your experience on our services.

\item \textbf{Adding Google Chrome, Google Chrome OS and Google Drive to
the Terms:} Our improved Terms now cover Google Chrome, Google
Chrome OS, and Google Drive, which also have service-specific terms
and policies to help you understand what’s unique to those services.

\item \textbf{No changes to our Privacy Policy:} We’re not making any changes to
the Google Privacy Policy and we haven't made any changes to the way
we treat your information. As a reminder, you can always visit your
Google Account to review your privacy settings and manage how your
data is used.
\end{itemize}

If you’re the guardian of a child under the age required to manage their own
Google Account and you use Family Link to manage their use of Google
services, please take some time to discuss these changes with them.

And of course, if you don’t agree to our new Terms and what we can expect from
each other as you use our services, you can find more information about your
options in our Frequently Asked Questions.

Thank you for using Google’s services.

\textbf{Your Google team}

\end{document}