### 2020-03-30 reviews/SIGBOVIK_2020_review_4.tex
reviews: Format Michael's unwitting review properly

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Michael Coblenz}
\rating{Objection}
\confidence{Your Honor}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
I’m afraid I must object to triple-blind review on the grounds that it, too, may be biased because the
reviewers still know they are reviewers. Have you considered quadruple-blind review, in which reviews are
elicited from reviewers without informing them that they are writing reviews?

\end{document}