### 2020-03-30 reviews/SIGBOVIK_2020_review_13.tex
Assign COVID-19 review

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Exponentiator}
\rating{Better and better every day, by leaps and bounds and\\\qquad\quad\; leaping leaps and bounding bounds}
\confidence{Those guys have faith, why shouldn't I?}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{36}
\papertitle{A thorough investigation of the degree to which the COVID-19 pandemic has enabled subpar-quality papers to make it into the proceedings of SIGBOVIK, by reducing the supply of authors willing to invest the necessary effort to produce high-quality papers}

\begin{document}

\maketitle

%%% Replace the following text with your review.
\vspace{-0.75in}
\center\includegraphics[height=4.25in]{walk-to-sky-exp}

\end{document}