### 2020-03-31 reviews/SIGBOVIK_2020_review_17.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{I got my friends to fill out a madlib}
\rating{47/69}
\confidence{16 micrometers}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

This paper discusses Oxycontin, with applications to ennui.  The paper begins
with a moribund demonstration of mental filtering, followed by examples of
solipsism, and concludes with a smelly repurposing of reptiles.

Overall, the presentation is glorious, with the small exception of page 291.34,
where the discussion of beer cans falls below expectations.  The author(s)
could perhaps merrily enslave this section with a brief enveloping of the
$\pi$/7th paragraph, but this would involve a reconsideration of the core
thesis.

The most surprising section, however, was on page 2, paragraph 666.  The
author(s) present a highly novel lazor, perhaps the first of it's kind.  Though
it seems like the graph (Fig. 512) is misprinted --- it appears to be missing
the R-coordinate?  Unclear.

Promising results, to be sure.  The reviewers are regretful of the impact this
will have on future research.

\end{document}