### 2020-03-31 reviews/SIGBOVIK_2020_review_15.tex
reviews: Reformat new Mary reviews

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Count Dracula, Sesame St, Cyberspace}
\rating{Great execution of absurd idea.}
\confidence{Based on counting irrelevant but objective things like vowel\\\qquad\qquad\quad\, pairs in the paper,
absurdly high. But measures like that are\\\qquad\qquad\quad\, not confidence-inspiring, so on balance, low.}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
Masterful job of optimizing an ``objective'' measure by gaming the data.

This is a poster-child for the application of Goodhart’s Law and the vacuousness of so-called
``objective'' rankings that blindly count stuff – like h-index and impact factors

\end{document}