### 2020-03-31 reviews/SIGBOVIK_2020_review_18.tex
reviews: Restore Internet searches to their former lack of glory

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{\textless ix@tchow.com\textgreater}
\rating{\textless sigbovik@gmail.com\textgreater}
\confidence{Re: Draft 2020 proceedings!}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{Review of Paper 31}
\papertitle{The SIGBOVIK 2020 field guide to plants}

\begin{document}

\maketitle

%%% Replace the following text with your review.
I think the review currently for the guide to plants should probably be attributed to the originator of the couplet (Edward Lear) rather than to "Prof. Jim McCann" -- I'm not sure there's much joke added (since it ended up on a paper referencing me) and I'm not really comfortable with something I didn't write being attributed to me.

-{}-Jim

\end{document}