### 2021-03-27 reviews/SIGBOVIK_2021_review_1.tex
all the ontime papers and some reviews

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2021 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Maya Harris, undergraduate student.}
\rating{6/10, 9/10 with rice.}
\confidence{100\%. Ok, fine. 95\%, since I know certainty freaks you out.}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

\subsection{Introduction}
Let me go! NO! You can't force me to read this! I already read enough on reddit. I don't care that you can't find any reviewers because of COVID-19. That's your problem (Shah, 2020). Let me GO! I said, let ME GO!

**muffled voices, struggling**

\subsection{Methods}
No one cares what your journal's impact factor is. What even is that? What do you expect me to say here? I barely listen in lectures and just wing all my homework. I don't even know what the difference between Windows and Linux is. I signed up for university to get a real job at FAANG or even Tesla so I can at least get my foot in the door in the valley, not help propagate your academic pyramid scheme. NO! I barely know how to read! Who cares about science or fun???? There's money to be made in industry. I want money.

**a door shuts**

**Rick Astley -- Never Gonna Give You Up starts to play on repeat in the closed room and the sound of water dripping echoes**

\textit{forty hours later}

\subsection{Results}
Ok, ok, ok. Just stop the music. Just, please, stop the music. I read the paper. I filled out all the sections from the template you gave me from autoreject.org, ok? I signed the thing that says I did this of my own free will and have no conflicts of interest.

\subsection{Discussion}
But you know, it actually really liked the article, so if you decide to accept -- what? You publish even papers even if reviewers recommend reject? How does that make any sense?

\subsection{Conclusion}
Ok whatever. Can I go now?

\subsection{References}

Shah, S. (2020). A Thorough Investigation of the Degree to which the COVID-19 Pandemic has Enabled Subpar-Quality Papers to Make it into the Proceedings of SIGBOVIK, by Reducing the Supply of Authors Willing to Invest the Necessary Effort to Produce High-Quality Papers. \textit{SIGBOVIK 2020}. URL: http://sigbovik.org/2020/proceedings.pdf

\end{document}