### 2021-03-27 reviews/SIGBOVIK_2021_review_2.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2021 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Hans-Peter Rüdisühli, Jasskönig vo de mittlere und vierte Schattegibeleggtäli}
\rating{Ahxeyt Zwöi, gmacht Driezäh, Difäränz Öuf!}
\confidence{Ha dänkt är loufi no is Näuu...}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
D'Chnäbäschterete vo dem Chlampf laht sich nid ring zämefasse, düecht mi aber es gröizigs Gspüder. Obwou hie wichtigi Erkenntnis us Schöppelimunggi, Houderebäsler, \& Totemügerli (1967) nid prantlet wärde, gstabets guet gfroorig u gäng no mappelig. Bsungers d'Hagletsche vo drobertrolige Spänkmügge wird vo buschtränggeligem Fodelebank sii u gürblet s'Bstotzige. Söu gäute!

\end{document}