### 2021-03-27 reviews/SIGBOVIK_2021_review_3.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2021 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Reviewer Number Two-and-a-half}
\rating{Excellent paper!}
\confidence{Absolutely bloody confident, as always!}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

%%% Replace the following text with your review.
\textbf{Excellent paper} -- truly marvellous! With a whiteness of 144, and an opacity of well over ninety percent, this paper promises excellent colour reproduction, which will be highly desirable for printing the inevitable graphs. Furthermore, this paper's grammage -- 80$g/m^2$ -- is fit for purpose, and keeps the price in check. Our proceedings need more paper of this calliper!

\end{document}