### 2021-03-27 reviews/SIGBOVIK_2021_review_4.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% SIGBOVIK 2020 TRIPLE BLIND REVIEW TEMPLATE
%%%
%%% Instructions:
%%%
%%% 1. Edit the author, rating, and confidence
%%%    fields.
%%% 2. Enter your review after the \maketitle
%%%    command
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{sigbovik-review}

%%%% Edit the following three lines.
%% To ensure the integrity of the triple-blind review
%% process, the contents of the \author field should not
%% reveal your identity.
\author{Reviewer from the future}
\rating{10}
\confidence{10}

%%%% The Proceedings Chair will fill in the following
%%%% two fields.
\papernum{NN}
\papertitle{UNKNOWN PAPER TITLE}

\begin{document}

\maketitle

I have come from the future to review this paper.
In the year 2035 time travel is invented. This paper is/was/will be the key to making it possible! At the time, I understand it was not terrifically well-received, due to its seeming lack of rigor, which admittedly did impede our understanding of its methods, likely causing a few years of delay. Thus, I have been assigned to make sure it gets accepted.

Unfortunately, this paper also indirectly resulted in an imminent grey goo scenario. Well, you win some, you lose some.

\end{document}