### 2021-03-27 papers.tex
listed all on-time papers w/ correct author names

\renewcommand{\thepage}{\roman{page}}
\setcounter{page}{3}

\newpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\addproceedings{Blindsight is also 2021}

\newcommand{\link}[2][]{\vspace{-1in} {\scriptsize #1 \url{#2}} \raisebox{-0.5\height}{\includegraphics[height=0.75in]{participation}}}

\addtrack
	{}{Big Dummy Track}
\addpaper
	{Refutation of the “Failure to remove the template text from your paper may result in your paper not being published” Conjecture}
	{Nicolas Hurtubise} %TODO: those other authors???
	{Keywords: component, formatting, style, styling, insert}
	{SIGBOVIK_2021_paper_1}
	{0cm}
	{}
\addpaper
	{Stop Doing Type Theory}
	{}
	{Keywords: Ranting, Raving, Muttering, Grumbling}
	{SIGBOVIK_2021_paper_2}
	{0cm}
	{}

\addpaper
	{``The SIGBOVIK paper to end all SIGBOVIK papers'' will not be appearing at this conference}
	{Thomas Chick}
	{Keywords: SIGBOVIK, Statistics, Finality}
	{SIGBOVIK_2021_paper_3}
	{0cm}
	{}
\addpaper
	{SIGBOVIK 2021 isn't named SIGCOVID}
	{Cameron Wong}
	{Keywords: covid, bovik, short}
	{SIGBOVIK_2021_paper_4}
	{0cm}
	{}
\addpaper
	{Back to Square One: Superhuman Performance in Chutes and Ladders Through Deep Neural Networks and Tree Search}
	{Dylan R. Ashley, Anssi Kanervisto and Brendan Bennett}
	{Keywords: Almost Monopoly, AlphaX, Artificial Neural Networks, Board Games, Deep Learning, Games With Boards, Machine Learning, Machine Learning That Matters, Reinforcement Learning, Tree Search}
	{SIGBOVIK_2021_paper_5}
	{0cm}
	{}
\addpaper
	{The Urinal Packing Problem in Higher Dimensions}
	{Shane Guan, Blair Chen, Skanda Kaashyap}
	{Keywords: PvsNP, Max\_Independent\_Set, TCS, Algorithms}
	{SIGBOVIK_2021_paper_6}
	{0cm}
	{}
\addpaper
	{How to get to second base and beyond - a constructive guide for mathematicians}
	{Raluca Jalaboi and Mads Eiler Hansen}
	{Keywords: damn, forgot, again, deadlines}
	{SIGBOVIK_2021_paper_7}
	{0cm}
	{}
\addpaper
	{Universal Insights with Multi-layered Embeddings}
	{Prophet \#1, Prophet \#2 and Prophet \#3}
	{Keywords: Machine learning, embedding, auto-encoders, Zalgo he comes, dimensional reduction}
	{SIGBOVIK_2021_paper_8}
	{0cm}
	{}
\addpaper
	{openCHEAT: Computationally Helped Error bar Approximation Tool - Kickstarting Science 4.0}
	{Bernhard Egger, Kevin Smith and Max Siegel} %TODO: put in crossed out David Cox?
	{Keywords: best, paper, award}
	{SIGBOVIK_2021_paper_9}
	{0cm}
	{}
%\addpaper
%	{AI is a Really, Really, Really Bad Idea}
%	{Molly de Blanc}
%	{Keywords: artificial intelligence, ethics, machine learning}
%	{SIGBOVIK_2021_paper_10}
%	{0cm}
%	{}
\addpaper
	{Revenge of the pith: Surveying the landscape of plant-powered scientific literature}
	{Vinay Uday Prabhu}
	{Keywords: Plant powered science, AI, Scientific method, Survey}
	{SIGBOVIK_2021_paper_11}
	{0cm}
	{}
\addpaper
	{Demystifying the Mortal Kombat Song}
	{J Devi and Chai-Tea Latte}
	{Keywords: mortal-kombat, truth, meaning-of-life}
	{SIGBOVIK_2021_paper_12}
	{0cm}
	{}
\addpaper
	{Unicode Magic Tricks}
	{Nicolas Hurtubise}
	{Keywords: Unicode, magic trick, emojis, bitwise operators, sleight of bits}
	{SIGBOVIK_2021_paper_13}
	{0cm}
	{}
\addpaper
	{Stone Tools as Palaeolithic Central Unit Processors}
	{}
	{Keywords: recursion, stone tools, computer science, handaxe, processor, prehistory, palaeogaming, psychedelic pragmatism}
	{SIGBOVIK_2021_paper_14}
	{0cm}
	{}
\addpaper
	{Winning the Rankings Game: A New, Wonderful, Truly Superior CS Ranking}
	{Diogenes}
	{Keywords: CSRankings, computer science rankings, bad proxies for quality, research quality, ego stroking, navel gazing, ranking algorithms}
	{SIGBOVIK_2021_paper_15}
	{0cm}
	{}
\addpaper
	{Solving reCAPTCHA v2 Using Deep Learning}
	{David Krajewski and Eugene Li}
	{Keywords: deep learning, recaptcha, automation, david}
	{SIGBOVIK_2021_paper_16}
	{0cm}
	{}
\addpaper
	{Deep Deterministic Policy Gradient Boosted Decision Trees}
	{Clayton Thorrez}
	{Keywords: machine learning, reinforcement learning, gradient, do these appear anywhere?, ``yes they do'': the proceedings chair}
	{SIGBOVIK_2021_paper_17}
	{0cm}
	{}
\addpaper
	{Video games in Fonts Fontemon}
	{Michael Mulet}
	{Keywords: font, video game, font video game, silly idea done seriously}
	{SIGBOVIK_2021_paper_18}
	{0cm}
	{}
\addpaper
	{Build your own 8-bit busy beaver on a breadboard!}
	{Robert Simmons}
	{Keywords: busy beaver, computability, turing machines, algorithms, breadboards, TTL, MacBook Pro}
	{SIGBOVIK_2021_paper_19}
	{0cm}
	{}
\addpaper
	{On Sigbovik Paper Maximization}
	{Josh Abrams}
	{Keywords: Maximization, Information Theory, Complexity, Divine Intervention, Academic Integrity, Big Small, Does Anybody Read These Things?, ``I Do'': The Proceedings Chair}
	{SIGBOVIK_2021_paper_20}
	{0cm}
	{}
\addpaper
	{Oracle Types}
	{Jason Reed and Akiva Leffert}
	{Keywords: types, databases, constraints, texting, machine translation, typescript}
	{SIGBOVIK_2021_paper_21}
	{0cm}
	{}
\addpaper
	{Instruction Programs}
	{James McCann}
	{Keywords: yoko ono, instruction programs, art jokes, one page papers}
	{SIGBOVIK_2021_paper_22}
	{0cm}
	{}
\addpaper
	{Inverted Code Theory: Manipulating Program Entropy}
	{Alan Hsu and Alex Xie}
	{Keywords: Tenet, Inverted, Temporal Pincer, Entropy, Turnstile, Inverted Code Theory, P vs. NP, Quantum Computing, Reinforcement Learning}
	{SIGBOVIK_2021_paper_23}
	{0cm}
	{}
\addpaper
	{Yet Another Lottery Ticket Hypothesis}
	{Aman Madaan and Gary Yao}
	{Keywords: lottery, random, language models}
	{SIGBOVIK_2021_paper_24}
	{0cm}
	{}
\addpaper
	{A Note on ``The Consent Hierarchy''}
	{Anonymous Nobody}
	{Keywords: freely given, reversible, informed, enthusiastic, specific}
	{SIGBOVIK_2021_paper_25}
	{0cm}
	{}
\addpaper
	{Lowestcase and uppestcase letters: Advances in derp learning}
	{Tom Murphy VII}
	{case analysis, derp learning, 3d manifold, chess, exponentiation, fonts}
	{SIGBOVIK_2021_paper_26}
	{0cm}
	{}

\addpaper
	{Soliterrible}
	{Sam Stern}
	{Keywords: solitaire, klondike, cards}
	{SIGBOVIK_2021_paper_27}
	{0cm}
	{}

\addpaper
	{Another Thorough Investigation of the Degree to which the COVID-19 Pandemic has Enabled Subpar-Quality Papers to Make it into SIGBOVIK, by Reducing the Supply of Authors Willing to Invest the Necessary Effort to Produce High-Quality Papers}
	{Shalin Shah}
	{Keywords: SIGBOVIK, COVID-19, Lazy, Low-Effort, Subpar-Quality}
	{SIGBOVIK_2021_paper_28}
	{0cm}
	{}
\addpaper
	{Spacecraft Attitude Determination and Control}
	{Freddie Rawlins}
	{Keywords: attitude, determination, happy satellites}
	{SIGBOVIK_2021_paper_29}
	{0cm}
	{}
\addpaper
	{A Complete Survey of 0-Dimensional Computer Graphics}
	{Hesper Yin, Oscar Dadfar, Max Slater, Anne He, Alice Lai, Emma Liu and Ryan Po}
	{Keywords: Computer Graphics, Geometry, Rendering, Animation}
	{SIGBOVIK_2021_paper_30}
	{0cm}
	{}
\addpaper
	{Macro-driven metalanguage for writing Pyramid Scheme programs}
	{Marcin Konowalczyk}
	{Keywords: syntax tree, pyramids, compilation, horizontal gene transfer, sorting, code golf}
	{SIGBOVIK_2021_paper_31}
	{0cm}
	{}
\addpaper
	{On the fundamental impossibility of refining the Theory of Everything by empirical observations: a computational theoretic perspective}
	{Zikuan Wang}
	{Keywords: Theory of Everything, Cellular Automaton, Emulation, the Universe}
	{SIGBOVIK_2021_paper_32}
	{0cm}
	{}
\addpaper
	{On the dire importance of MRU caches for human survival (against Skynet)}
	{Darío de la Fuente García, Félix Áxel Gimeno Gil, Juan Carlos Morales Vega and Borja Rodríguez Gálvez}
	{Keywords: Skynet, MS paint, cache, MRU, dMRU, sMRU, NEP}
	{SIGBOVIK_2021_paper_33}
	{0cm}
	{}
\addpaper
	{Dependent Stringly-Typed Programming}
	{G. Allais}
	{Keywords: dependent stringly typed programming, best practices, agda}
	{SIGBOVIK_2021_paper_34}
	{0cm}
	{}
\addpaper
	{Tensorflow for Abacus Processing Units}
	{Robert McCraith}
	{Keywords: machine learning, tensor flow, abacus, mathematics, calculus, differentiation, computation}
	{SIGBOVIK_2021_paper_35}
	{0cm}
	{}
\addpaper
	{Critical Investigations on Avians: Surveillance, Computational Amorosities, and Machines}
	{Brandon Bohrer and Connie Chau}
	{Keywords: Birds, Not, Real}
	{SIGBOVIK_2021_paper_36}
	{0cm}
	{}
\addpaper
	{NetPlop: A moderately-featured presentation editor built in NetLogo}
	{Patrick Steinmann}
	{Keywords: agent-based slideshows, slide-based models, singularity}
	{SIGBOVIK_2021_paper_37}
	{0cm}
	{}
\addpaper
	{RadicAI: A Radical, Though Not Entirely New, Approach to AI Paper Naming}
	{James McCann and Mike McCann}
	{Keywords: lottery, random, language models}
	{SIGBOVIK_2021_paper_38}
	{0cm}
	{}
\addpaper
	{Story Time}
	{James McCann and Mike McCann}
	{Keywords: story time, time for stories, the time of stories, stories? it's time, our time is now our time is stories, big data? I get it.}
	{SIGBOVIK_2021_paper_39}
	{0cm}
	{}
\addpaper
	{If It Type-checks, It Works: FoolProof Types As Specifications}
	{Brandon Wu}
	{Keywords: PL, languages, types, specifications, correctness, type-checking, functional programming}
	{SIGBOVIK_2021_paper_40}
	{0cm}
	{}
\addpaper
	{The Newcomb-Benford Law, Applied to Binary Data: An Empirical and Theoretic Analysis}
	{Gabriel Chuang and Brandon Wu}
	{Keywords: Numerical Methods, Verification, Poor Proof Style, Statistics}
	{SIGBOVIK_2021_paper_41}
	{0cm}
	{}
\addpaper
	{What Lothar Collatz Thinks of the CMU Computer Science Curriculum}
	{Gabriel Chuang and Brandon Wu}
	{Pedagogy, Precognition, Theoretical Computer Science, Coincidences, Sequences, Category Theory, Computability}
	{SIGBOVIK_2021_paper_42}
	{0cm}
	{}
\addpaper
	{On the Origin of Species of Self-Supervised Learning}
	{Samuel Albanie, Erika Lu and Joao Henriques}
	{Keywords: self-supervised learning, origin of species, artificial naturalism}
	{SIGBOVIK_2021_paper_43}
	{0cm}
	{}
\addpaper
	{Opening Moves in 1830: Strategy in Resolving the N-way Prisoner’s Dilemma}
	{Philihp Busby and Daniel Ribeiro E Sousa}
	{Keywords: boardgame, opening, strategy, deterministic, auction}
	{SIGBOVIK_2021_paper_44}
	{0cm}
	{}

\addreview
	{SIGBOVIK_2021_review_1}
	{0cm}