### 2021-04-01 papers.tex
add badge PDFs, and badge to last paper

\renewcommand{\thepage}{\roman{page}}
\setcounter{page}{3}

\newpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\addproceedings{Blindsight is also 2021}

\newcommand{\link}[2][]{\vspace{-1in} {\scriptsize #1 \url{#2}} \raisebox{-0.5\height}{\includegraphics[height=0.75in]{participation}}}
\newcommand{\badge}[1]{\includegraphics[height=0.75in]{badges/badge_#1}}
\newcommand{\badges}[1]{\vspace{-1in} \raisebox{-0.5\height}{#1}}

\addtrack
    {}{Fun(?) and Games Track}
\addpaper
    {Back to Square One: Superhuman Performance in Chutes and Ladders Through Deep Neural Networks and Tree Search}
    {Dylan R. Ashley, Anssi Kanervisto and Brendan Bennett}
    {Keywords: Almost Monopoly, AlphaX, Artificial Neural Networks, Board Games, Deep Learning, Games With Boards, Machine Learning, Machine Learning That Matters, Reinforcement Learning, Tree Search}
    {SIGBOVIK_2021_paper_5}
    {0cm}
    {\badges{\badge{unusable}}}
\addpaper
    {Demystifying the Mortal Kombat Song}
    {J Devi and Chai-Tea Latte}
    {Keywords: mortal-kombat, truth, meaning-of-life}
    {SIGBOVIK_2021_paper_12}
    {0cm}
    {}
\addpaper
    {Unicode Magic Tricks}
    {Nicolas Hurtubise}
    {Keywords: Unicode, magic trick, emojis, bitwise operators, sleight of bits}
    {SIGBOVIK_2021_paper_13}
    {0cm}
    {}
\addpaper
    {Video games in Fonts Fontemon}
    {Michael Mulet}
    {Keywords: font, video game, font video game, silly idea done seriously}
    {SIGBOVIK_2021_paper_18}
    {0cm}
    {}
\addreview
    {SIGBOVIK_2021_review_5}
    {0cm}

\addpaper
    {Soliterrible}
    {Sam Stern}
    {Keywords: solitaire, klondike, cards}
    {SIGBOVIK_2021_paper_27}
    {0cm}
    {\badges{\badge{javascript}}}
\addpaper
    {Opening Moves in 1830: Strategy in Resolving the N-way Prisoner’s Dilemma}
    {Philihp Busby and Daniel Ribeiro E Sousa}
    {Keywords: boardgame, opening, strategy, deterministic, auction}
    {SIGBOVIK_2021_paper_44}
    {0cm}
    {}

\addtrack
    {}{Obligatory Machine Learning Track}
\addpaper
    {Universal Insights with Multi-layered Embeddings}
    {Prophet \#1, Prophet \#2 and Prophet \#3}
    {Keywords: Machine learning, embedding, auto-encoders, Zalgo he comes, dimensional reduction}
    {SIGBOVIK_2021_paper_8}
    {0cm}
    {}
\addpaper
    {Solving reCAPTCHA v2 Using Deep Learning}
    {David Krajewski and Eugene Li}
    {Keywords: deep learning, recaptcha, automation, david}
    {SIGBOVIK_2021_paper_16}
    {0cm}
    {}
\addpaper
    {Deep Deterministic Policy Gradient Boosted Decision Trees}
    {Clayton W. Thorrez}
    {Keywords: machine learning, reinforcement learning, gradient, do these appear anywhere?, ``yes they do'': the proceedings chair}
    {SIGBOVIK_2021_paper_17}
    {0cm}
    {}
\addpaper
    {Tensorflow for Abacus Processing Units}
    {Robert McCraith}
    {Keywords: machine learning, tensor flow, abacus, mathematics, calculus, differentiation, computation}
    {SIGBOVIK_2021_paper_35}
    {0cm}
    {}

\addpaper
    {RadicAI: A Radical, Though Not Entirely New, Approach to AI Paper Naming}
    {Jim McCann and Mike McCann}
    {Keywords: lottery, random, language models}
    {SIGBOVIK_2021_paper_38}
    {0cm}
    {}

\addtrack
    {}{Followup Track}
\addpaper
    {A Note on ``The Consent Hierarchy''}
    {}
    {Keywords: freely given, reversible, informed, enthusiastic, specific}
    {SIGBOVIK_2021_paper_25}
    {0cm}
    {}

\addpaper
    {Another Thorough Investigation of the Degree to which the COVID-19 Pandemic has Enabled Subpar-Quality Papers to Make it into SIGBOVIK, by Reducing the Supply of Authors Willing to Invest the Necessary Effort to Produce High-Quality Papers}
    {Shalin Shah}
    {Keywords: SIGBOVIK, COVID-19, Lazy, Low-Effort, Subpar-Quality}
    {SIGBOVIK_2021_paper_28}
    {0cm}
    {}

\addpaper
    {Story Time}
    {Jim McCann and Mike McCann}
    {Keywords: story time, time for stories, the time of stories, stories? it's time, our time is now our time is stories, big data? I get it.}
    {SIGBOVIK_2021_paper_39}
    {0cm}
    {}

\addtrack
    {}{``Type'' Track}
\addpaper
    {Stop Doing Type Theory}
    {}
    {Keywords: Ranting, Raving, Muttering, Grumbling}
    {SIGBOVIK_2021_paper_2}
    {0cm}
    {}
\addreview
    {SIGBOVIK_2021_review_3}
    {0cm}
\addpaper
    {If It Type-checks, It Works: FoolProof Types As Specifications}
    {Brandon Wu}
    {Keywords: PL, languages, types, specifications, correctness, type-checking, functional programming}
    {SIGBOVIK_2021_paper_40}
    {0cm}
    {}
\addpaper
    {Oracle Types}
    {Akiva Leffert and Jason Reed}
    {Keywords: types, databases, constraints, texting, machine translation, typescript}
    {SIGBOVIK_2021_paper_21}
    {0cm}
    {\badges{\badge{unavailable}}}
\addpaper
    {Lowestcase and uppestcase letters: Advances in derp learning}
    {Tom Murphy VII}
    {case analysis, derp learning, 3d manifold, chess, exponentiation, fonts}
    {SIGBOVIK_2021_paper_26}
    {0cm}
    {}
\addpaper
    {Dependent Stringly-Typed Programming}
    {gallais}
    {Keywords: dependent stringly typed programming, best practices, agda}
    {SIGBOVIK_2021_paper_34}
    {0cm}
    {\badges{\badge{literate}}}
\addpaper
    {Yet Another Lottery Ticket Hypothesis}
    {Aman Madaan and Gary Yao}
    {Keywords: lottery, random, language models}
    {SIGBOVIK_2021_paper_24}
    {0cm}
    {}

\addtrack
    {}{(Psycho)metrics Track}
\addpaper
    {Spacecraft Attitude Determination and Control}
    {Freddie Rawlins}
    {Keywords: attitude, determination, happy satellites}
    {SIGBOVIK_2021_paper_29}
    {0cm}
    {}

\addpaper
    {Instruction Programs}
    {Jim McCann}
    {Keywords: yoko ono, instruction programs, art jokes, one page papers}
    {SIGBOVIK_2021_paper_22}
    {0cm}
    {}

\addpaper
    {Winning the Rankings Game: A New, Wonderful, Truly Superior CS Ranking}
    {Diogenes}
    {Keywords: CSRankings, computer science rankings, bad proxies for quality, research quality, ego stroking, navel gazing, ranking algorithms}
    {SIGBOVIK_2021_paper_15}
    {0cm}
    {}

\addpaper
    {openCHEAT: Computationally Helped Error bar Approximation Tool - Kickstarting Science 4.0}
    {Bernhard Egger, Kevin Smith and Max Siegel} %TODO: put in crossed out David Cox?
    {Keywords: best, paper, award}
    {SIGBOVIK_2021_paper_9}
    {0cm}
    {}

\addreview
    {SIGBOVIK_2021_review_1}
    {0cm}

\addpaper
    {On the dire importance of MRU caches for human survival (against Skynet)}
    {Darío de la Fuente García, Félix Áxel Gimeno Gil, Juan Carlos Morales Vega and Borja Rodríguez Gálvez}
    {Keywords: Skynet, MS paint, cache, MRU, dMRU, sMRU, NEP}
    {SIGBOVIK_2021_paper_33}
    {0cm}
    {}

\addtrack
    {}{Not Really Biology But Closer to it Than the Other Papers Track}
\addpaper
    {Revenge of the pith: Surveying the landscape of plant-powered scientific literature}
    {Vinay Uday Prabhu}
    {Keywords: Plant powered science, AI, Scientific method, Survey}
    {SIGBOVIK_2021_paper_11}
    {0cm}
    {}

\addreview
    {SIGBOVIK_2021_review_4}
    {0cm}

\addpaper
    {On the Origin of Species of Self-Supervised Learning}
    {Samuel Albanie, Erika Lu and João Henriques}
    {Keywords: self-supervised learning, origin of species, artificial naturalism}
    {SIGBOVIK_2021_paper_43}
    {0cm}
    {}
\addpaper
    {Critical Investigations on Avians: Surveillance, Computational Amorosities, and Machines}
    {Brandon Bohrer and Connie Chau}
    {Keywords: Birds, Not, Real}
    {SIGBOVIK_2021_paper_36}
    {0cm}
    {}

\addpaper
    {The Urinal Packing Problem in Higher Dimensions}
    {Shane Guan, Blair Chen, Skanda Kaashyap}
    {Keywords: PvsNP, Max\_Independent\_Set, TCS, Algorithms}
    {SIGBOVIK_2021_paper_6}
    {0cm}
    {}

\addtrack
    {}{ApPLied Theory}
\addpaper
    {The Newcomb-Benford Law, Applied to Binary Data: An Empirical and Theoretic Analysis}
    {Gabriel Chuang and Brandon Wu}
    {Keywords: Numerical Methods, Verification, Poor Proof Style, Statistics}
    {SIGBOVIK_2021_paper_41}
    {0cm}
    {}
\addpaper
    {How to get to second base and beyond - a constructive guide for mathematicians}
    {Raluca Jalaboi and Mads Eiler Hansen}
    {Keywords: damn, forgot, again, deadlines}
    {SIGBOVIK_2021_paper_7}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addpaper
    {NetPlop: A moderately-featured presentation editor built in NetLogo}
    {Patrick Steinmann}
    {Keywords: agent-based slideshows, slide-based models, singularity}
    {SIGBOVIK_2021_paper_37}
    {0cm}
    {\badges{\badge{netlogo}}}

\addtrack{}{(Meta)physics}
\addpaper
    {A Complete Survey of 0-Dimensional Computer Graphics}
    {Hesper Yin, Oscar Dadfar, Max Slater, Anne He, Alice Lai, Emma Liu and Po Po}
    {Keywords: Computer Graphics, Geometry, Rendering, Animation}
    {SIGBOVIK_2021_paper_30}
    {0cm}
    {}
\addpaper
    {Macro-driven metalanguage for writing Pyramid Scheme programs}
    {Marcin Konowalczyk}
    {Keywords: syntax tree, pyramids, compilation, horizontal gene transfer, sorting, code golf}
    {SIGBOVIK_2021_paper_31}
    {0cm}
    {\badges{\hspace{0.5in}\badge{mlm}}}
\addpaper
    {On the fundamental impossibility of refining the Theory of Everything by empirical observations: a computational theoretic perspective}
    {Zikuan Wang}
    {Keywords: Theory of Everything, Cellular Automaton, Emulation, the Universe}
    {SIGBOVIK_2021_paper_32}
    {0cm}
    {}
\addpaper
    {Inverted Code Theory: Manipulating Program Entropy}
    {usH nalA dna eiX xelA}
    {Keywords: Tenet, Inverted, Temporal Pincer, Entropy, Turnstile, Inverted Code Theory, P vs. NP, Quantum Computing, Reinforcement Learning}
    {SIGBOVIK_2021_paper_23}
    {0cm}
    {}
\addreview
    {SIGBOVIK_2021_review_2}
    {0cm}

\addtrack
    {}{Definitely Finite Track}

%\addpaper
%    {AI is a Really, Really, Really Bad Idea}
%    {Molly de Blanc}
%    {Keywords: artificial intelligence, ethics, machine learning}
%    {SIGBOVIK_2021_paper_10}
%    {0cm}
%    {}

\addpaper
    {Stone Tools as Palaeolithic Central Unit Processors}
    {}
    {Keywords: recursion, stone tools, computer science, handaxe, processor, prehistory, palaeogaming, psychedelic pragmatism}
    {SIGBOVIK_2021_paper_14}
    {0cm}
    {}

\addpaper
    {Build your own 8-bit busy beaver on a breadboard!}
    {Robert J. Simmons}
    {Keywords: busy beaver, computability, turing machines, algorithms, breadboards, TTL, MacBook Pro}
    {SIGBOVIK_2021_paper_19}
    {0cm}
    {}

\addpaper
    {What Lothar Collatz Thinks of the CMU Computer Science Curriculum}
    {Gabriel Chuang and Brandon Wu}
    {Pedagogy, Precognition, Theoretical Computer Science, Coincidences, Sequences, Category Theory, Computability}
    {SIGBOVIK_2021_paper_42}
    {0cm}
    {}

\addtrack
    {}{Recursive Track}
\addpaper
    {On Sigbovik Paper Maximization}
    {Josh Abrams}
    {Keywords: Maximization, Information Theory, Complexity, Divine Intervention, Academic Integrity, Big Small, Does Anybody Read These Things?, ``I Do'': The Proceedings Chair}
    {SIGBOVIK_2021_paper_20}
    {0cm}
    {}

\addpaper
    {SIGBOVIK 2021 isn't named SIGCOVID}
    {Cameron Wong}
    {Keywords: covid, bovik, short}
    {SIGBOVIK_2021_paper_4}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addpaper
    {Refutation of the “Failure to remove the template text from your paper may result in your paper not being published” Conjecture}
    {Nicolas Hurtubise} %TODO: those other authors???
    {Keywords: component, formatting, style, styling, insert}
    {SIGBOVIK_2021_paper_1}
    {0cm}
    {}

\addpaper
    {``The SIGBOVIK paper to end all SIGBOVIK papers'' will not be appearing at this conference}
    {Thomas Chick}
    {Keywords: SIGBOVIK, Statistics, Finality}
    {SIGBOVIK_2021_paper_3}
    {0cm}
    {}