### 2022-04-08 papers.tex
Final 2022 version??? Probably not

\renewcommand{\thepage}{\roman{page}}
\setcounter{page}{3}

\newpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\addproceedings{TODO: Tell next year's committee that they should probably avoid further milking the blindsight joke}

\newcommand{\link}[2][]{\vspace{-1in} {\scriptsize #1 \url{#2}} \raisebox{-0.5\height}{\includegraphics[height=0.75in]{participation}}}
\newcommand{\badge}[1]{\includegraphics[height=0.75in]{badges/badge_#1}}
\newcommand{\badges}[1]{\vspace{-1in} \raisebox{-0.5\height}{#1}}

\addtrack
    {}{Programming Languages}
\addpaper
    {Modernized Python}
    {Daniel Ng}
    {Keywords:programming languages, type theory, python, sml}
    {SIGBOVIK_2022_paper_0413}
    {0cm}
    {}
\addpaper
    {Grand Challenges in Programming Languages Position Paper: What, if anything, does multiplication even mean?}
    {Jim McCann}
    {multiplication, punctuation: *, punctuation: x, punctuation: ., grand challenges}
    {SIGBOVIK_2022_paper_3280}
    {0cm}
    {}
\addpaper
    {A Type-and-Affect System for Semisignificant Whitespace}
    {Stefan Muller}
    {Keywords: whitespace, white space, white\\nspace}
    {SIGBOVIK_2022_paper_6424}
    {0cm}
    {}

\addpaper
    {brainfuck++: a much needed extension to brainfuck}
    {Cédric Ho Thanh}
    {Keywords: Programing langage, brainfuck, concurrency, machine learning}
    {SIGBOVIK_2022_paper_3995}
    {0cm}
    {\badges{\badge{pagenumbers}}}
%\addreview
%    {SIGBOVIK_2021_review_5}
%    {0cm}

\addtrack
    {}{Other Languages}

\addpaper
    {Tironiculum: Latin Speech Recognition via Latin Text-to-Speech}
    {Lee Butterman}
    {Keywords: Latin, Automatic speech recognition, AI, Machine Learning, Abundant citations, Life advice, State of the Art, Competition winner, Metagaming a competition}
    {SIGBOVIK_2022_paper_0950}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addpaper
    {Abecedarial Acrostic, Alphabetized Amusingly Because Beings Blissfully Cause Celebratory Centennials... (Note: Full Title is Longer)}
    {Jacob Weiner}
    {Keywords: Records, Useless, Alphabet, Gibberish}
    {SIGBOVIK_2022_paper_1598}
    {0cm}
    {}

\addpaper
    {On ``Ra-men, Ra-men ramen ramen''}
    {LAPP Lab}
    {Keywords: language, ramen, space-time continuum}
    {SIGBOVIK_2022_paper_8333}
    {0cm}
    {}

\addpaper
    {ACTION: A Catchy Title Is all yOu Need!}
    {Bernhard Egger, Kevin Smith, Thomas O'Connell and Max Siegel}
    {Keywords: catchy, title, all you need}
    {SIGBOVIK_2022_paper_8346}
    {0cm}
    {}

\addpaper
    {A Deep Learning Approach for Deeply Inaccurate Wordle Solving}
    {Ahana Deb and Sayan Goswami}
    {Keywords: wordle, ordlew, rdlewo, dlewor, leword, ewordl}
    {SIGBOVIK_2022_paper_7404}
    {0cm}
    {}

\addtrack
    {}{Systems}

\addpaper
    {Edward, edMUnD \& Edwin: Line-Based Text Editing for the 21st Century}
    {Natalia Posting and Katie Wolfe}
    {Keywords: line-based text editing, line-cringe text editing, synergy}
    {SIGBOVIK_2022_paper_6729}
    {0cm}
    {}
\addpaper
    {Redundant Coupling}
    {Peter Kos}
    {Keywords: software engineering, coupling, cohesion, rust, martin fowler, fowler, james gosling, donald knuth}
    {SIGBOVIK_2022_paper_8888}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {A Free Computer Vision Lesson for Car Manufacturers or It is Time to Retire the Erlkönig}
    {Maximilian Weiherer and Bernhard Egger}
    {Keywords: awesome paper, great work, mind blowing results}
    {SIGBOVIK_2022_paper_9466}
    {0cm}
    {}

\addtrack
    {}{Theory}
\addpaper
    {Destructive Logic}
    {Runming Li}
    {Keywords: logic, constructive logic, destructive logic, type theory}
    {SIGBOVIK_2022_paper_6801}
    {0cm}
    {}
\addpaper
    {Overlap-Maximal Graph Labelings: Graph Labelings with Non-Disjoint Vertex and Edge Sets (and how they can be used for encryption, poetry, and breaking mathematics)}
    {Gabriel Chuang}
    {Keywords: Graph Theory, Cryptography, Fox in Socks by Dr. Seuss}
    {SIGBOVIK_2022_paper_6809}
    {0cm}
    {}

\addpaper
    {Neo-classical Logic: Une Logique non classique}
    {Martin Vassor and Fangyi Zhou}
    {Keywords: Logic, Meta-Metatheory, Law of Excluded Middle, Law of Contradiction, Decidability, Macron}
    {SIGBOVIK_2022_paper_2522}
    {0cm}
    {}

\addpaper
    {A Patriotic Analysis of Programming Paradigms}
    {Jacob Weiner}
    {Keywords: Patriotism, Constitution, Dependent Type Theory}
    {SIGBOVIK_2022_paper_1689}
    {0cm}
    {}

\addpaper
    {On Ruinment: Ruination Theory and its Consequents}
    {Luna Tascheter}
    {Ruination Theory, Ruinment, T0T4L PWN4G∃, Category Theory, Adam Conover, Adam Sandler, Hallmark, Waste of Paper, Orange}
    {SIGBOVIK_2022_paper_6808}
    {0cm}
    {}

\addtrack
    {}{Astrophysics}
\addpaper
    {Method and Tool for Estimating the Mass of the Black Hole Located in the Office of Immigration, Refugees and Citizenship Canada Causing a Supermassive Time Dilation in the Visa Extension Process}
    {Étienne Trottoir-Barré, Richard von Pamplemousse and Jessica G. Lasso}
    {Keywords: black hole, abracadabra, hyper driven devices, natural language processing, Chevy Tahoes}
    {SIGBOVIK_2022_paper_0714}
    {0cm}
    {}
\addpaper
    {Black Hole Computation}
    {Matias Scharager}
    {Keywords: monad, computation, compilation, computational trinitarianism, parallel universe, parallel computing}
    {SIGBOVIK_2022_paper_2805}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {Schrödinger’s SAT: Generalizing Quantum Bogosort to Prove P = NP Under Many-Worlds Quantum Mechanics}
    {Melody Horn}
    {Keywords: satisfiability, time complexity, revolutions in our understanding of computing, unsolved problems in millennium prize eligibility}
    {SIGBOVIK_2022_paper_6784}
    {0cm}
    {}

\addpaper
    {Solving Double Execution of Java’s paint() Method by Counting Down to the Heat Death of the Universe (plus language compendium)}
    {Braden Oh, Vedaant Kuchhal, Junseok Kang, Andrew Mascillaro, Justin Kunimune, Shashank Swaminathan, Devlin Ih, Elias Gabriel, Audrey Lee, Colin Snow, Hyunkyung Rho, Ben Morris, Solomon Greenberg, Mahima Beltur, Anusha Datar, Jonathan Kelley, Pranavi Boyalakuntla, Xander Hughes, Devyn Oh, Aidan Schmitigal}
    {Keywords: Efficient, Effective, Compendium, Languages, Heat death, Overkill}
    {SIGBOVIK_2022_paper_1777}
    {0cm}
    {}
\addtrack
    {}{Climate Science}

\addpaper
    {Towards Cloud Computing}
    {Alex Xie and Alan Hsu}
    {Keywords: cloud computing, deep learning, machine learning, computer vision, toddlers, kahoot, clout, drip, label smoothing, data science, animals, atmospheric sciences, meteorology}
    {SIGBOVIK_2022_paper_2551}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addpaper
    {Ecological Memory Management: Beyond Garbage Collection}
    {Erik Derohanian, Dann Toliver and Saul Field}
    {Keywords: Garbage Collection, Compost, Memory Management, Organic Computing}
    {SIGBOVIK_2022_paper_2558}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addreview
    {SIGBOVIK_2022_review_2}
    {0cm}

\addpaper
    {Infrastructure-as-PowerPoint: A No-Code Approach to Cloud Infrastructure Management}
    {Tobias Pfandzelter}
    {Keywords: cloud computing, powerpoint, slideshow bob}
    {SIGBOVIK_2022_paper_3392}
    {0cm}
    {}

\addpaper
    {On the Possibilities and Challenges of Organic UAV-Assisted MEC}
    {S. Wallow, Cardi Nalle, Robin, P. Cock, and Sky Lark}
    {Keywords:chick-a-dee, caw caw, chirp chirp, hoot hoot}
    {SIGBOVIK_2022_paper_9501}
    {0cm}
    {}

\addtrack{}{A Brief Musical Interlude}
%\fancypagestyle{fancypyramid}{\fancyheadoffset[roh]{0.75in}}
\addpaper
    {Baby Sharks are More than Sharks . They are Earworms Sung More than Happy Birthday Perhaps}
    {Kofi Oduro}
    {Keywords: Sharks, Happy Birthday, Swimming Shark Problem, Shark Week, billions, humans, kindergarten, PRe-K}
    {SIGBOVIK_2022_paper_6744}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {Everybody Clap Your Hands: The Cha Cha Slide is Turing Complete}
    {Harrison Goldstein}
    {Keywords: Turing complete, Cha Cha Slide, line dance, party, complexity}
    {SIGBOVIK_2022_paper_5708}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {Exhaustive Survey of Rickrolling in Academic Literature}
    {Benoit Baudry and Martin Monperrus}
    {Keywords: Rickrolling, Academia, Rick Astley}
    {SIGBOVIK_2022_paper_8755}
    {0cm}
    {}

\addtrack
    {}{Best Practices}
%\addpaper
%    {Investigations and Difficulties in Real - to -Sim}
%    {Sashank Tirumala}
%    {Keywords: Robotics, Machine Learning, Deep Learning, Reinforcement Learning}
%    {SIGBOVIK_2022_paper_3472}
%    {0cm}
%    {}
\addpaper
    {Quadruple-Blind Peer Review}
    {Yuriko Shu}
    {Keywords: Peer Review, Quadruple Blind Reviewing, Bias Reduction}
    {SIGBOVIK_2022_paper_3958}
    {0cm}
    {}
\addreview
    {SIGBOVIK_2022_review_3}
    {0cm}

\addpaper
    {Optimal degeneracy through OwO based variable names}
    {Fwans Skawman}
    {Keywords: OwO, What's, This}
    {SIGBOVIK_2022_paper_4751}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {Multiplication by repeated addition, with fraction handling.}
    {Jim McCann}
    {Keywords: multiplication, punctuation: *, C++, source code}
    {SIGBOVIK_2022_paper_1824}
    {0cm}
    {}
\addpaper
    {Objective Correlation Metrics for Quality of Code Estimation}
    {Eleftheria Chatziargyriou and Konstantinos Kanavouras}
    {Keywords: programming, quality, code, objective}
    {SIGBOVIK_2022_paper_1957}
    {0cm}
    {}

\addtrack
    {}{Algorithmic Advances}

\addpaper
    {Functorial wrappers for high-dimensional classification algorithms}
    {James Carzon}
    {Keywords: forgetful functor, classification, high dimensional data}
    {SIGBOVIK_2022_paper_7603}
    {0cm}
    {}

\addpaper
    {(Un)helper functions}
    {Kevin Smith and Bernhard Egger}
    {Keywords: algorithms, algorithm patent trolling, cite us please!}
    {SIGBOVIK_2022_paper_2133}
    {0cm}
    {}
\addpaper
    {A sometimes-accurate O(1) search algorithm}
    {A College Freshman}
    {Keywords:fast, search, search algorithm, searching algorithm, searching, how many ways can i say search algorithm}
    {SIGBOVIK_2022_paper_4297}
    {0cm}
    {}

\addtrack
    {}{Mathematic Retreats}

\addpaper
    {Improved Data and Instruction Locality in Long Division}
    {Isaac Grosof, Isaac Grosof}
    {Keywords: performance, division, locality, data, instruction}
    {SIGBOVIK_2022_paper_0463}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {The New New Math: using sentiment analysis of mathematics word problems to gauge children's reactions to teaching 8-bit floating point arithmetic for the new California public school math curriculum}
    {Lee Butterman}
    {Keywords: education, early childhood education, won't someone think of the children, won't someone think of the chiiiiiiildren, new math, California math education guidelines, data science, calculus, graphics processing units, dozens of petaflops, sentiment analysis, numerology, fear of Arabic numerals, 8-bit floating point}
    {SIGBOVIK_2022_paper_4483}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {Infix Modifiers for Flexible Multiplication}
    {Jim McCann}
    {Keywords: multiplication, punctuation: *, infix, IMMs}
    {SIGBOVIK_2022_paper_4549}
    {0cm}
    {}

\addtrack
    {}{Aesthetics}

\addpaper
    {Real-Time Foliage Simulation}
    {Emma Liu, Sanjay Salem, Daniel Zeng, Mia Tang, Hesper Yin, George Ralph, Anne He, Max Slater}
    {Keywords: real-time, simulation, graphics, rendering}
    {SIGBOVIK_2022_paper_0683}
    {0cm}
    {}

\addpaper
    {Using deep CNNs to prove that I look better than Tom Cruise and Shah Rukh Khan combined}
    {Sagar Bharadwaj Kalasibail Seetharam}
    {Keywords: CNN, Machine Learning, Image Compression}
    {SIGBOVIK_2022_paper_6854}
    {0cm}
    {}

\addpaper
    {EBMP: Efficient Bitmap Encodings on Ethereum Virtual Machines}
    {0XMOSTIMA, Yunsong Liu and Peiyuan Liao}
    {Keywords: still image coding, blockchain, smart contracts, non fungible tokens}
    {SIGBOVIK_2022_paper_5649}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addpaper
    {A Machine Learning Approach To Classifying Cuteness}
    {Anoushka Shrivastava}
    {Keywords: machine learning, classification, cuteness}
    {SIGBOVIK_2022_paper_9240}
    {0cm}
    {}

\addpaper
    {Attractiveness Learning: A General Solution for the Cold-Start Problem}
    {Yajuan Gu and Yuxun Lu}
    {neural network, attractive mechanism, recommender system, cold-start problem}
    {SIGBOVIK_2022_paper_2018}
    {0cm}
    {}
\addtrack
    {}{Hardware}

\addpaper
    {Correct-It-Yourself Paper Updater}
    {Thomas Chick}
    {Keywords: corrections, arts and crafts, reader participation, sacrilege? sure}
    {SIGBOVIK_2022_paper_4967}
    {0cm}
    {}

\addpaper
    {Harder Drive: Hard drives we didn't want or need}
    {Tom Murphy}
    {Keywords: hard drive, block device, ipv4, tetris, covid-19, chainsaws}
    {SIGBOVIK_2022_paper_4275}
    {0cm}
    {}
\addpaper
    {Just a Regular Paper}
    {Jason Wang}
    {Keywords: Money, Finance, Banking, Social Issues, Machine Learning, Ohio}
    {SIGBOVIK_2022_paper_2759}
    {0cm}
    {}

\addpaper
    {Wearable RF-shield repositories}
    {Abe Wits and J. Mijn}
    {Keywords: security, wifi, cable}
    {SIGBOVIK_2022_paper_9009}
    {0cm}
    {\badges{\badge{pagenumbers}}}
\addtrack
    {}{Black Boxes}

\addpaper
    {A 23 MW data centre is all you need}
    {Samuel Albanie, Dylan Campbell and João Henriques}
    {Keywords: machine learning, prophecy, Molly}
    {SIGBOVIK_2022_paper_4719}
    {0cm}
    {}

\addpaper
    {When Pull Comes To Shove... Do Both!}
    {Ullas A}
    {Keywords: Parallel Programming, Graphs, Connected Components, GPUs}
    {SIGBOVIK_2022_paper_8597}
    {0cm}
    {}

\addpaper
    {man exorcism}
    {}
%    {Akiva Leffert}
    {Keywords:exorcisms, unix, demonology}
    {SIGBOVIK_2022_paper_9978}
    {0cm}
    {}

\addpaper
    {2x a fake submission}
    {Tim Toady}
    {Keywords: one letter keywords are not accepted: f, 21 is not a correct keyword: it contains no letters, =42 is not a correct keyword: it contains no letters}
    {SIGBOVIK_2022_paper_8085}
    {0cm}
    {}

\addreview
    {SIGBOVIK_2022_review_1}
    {0cm}
\addtrack{}{SIGBOVIK}

\addpaper
    {Analysis of A New Error Calculation Technique}
    {Jevin Tong} 
    {Keywords: late, erroneous, hackneyed}
    {SIGBOVIK_2022_paper_7011}
    {0cm}
    {\badges{\badge{pagenumbers}}}

\addpaper
    {A Third Thorough Investigation of the Degree to which the COVID-19 Pandemic has Enabled Subpar-Quality Papers to Make it into SIGBOVIK, by Reducing the Supply of Authors Willing to Invest the Necessary Effort to Produce High-Quality Papers}
    {Shalin Shah}
    {Keywords: SIGBOVIK, COVID-19, Lazy, Low-Effort}
    {SIGBOVIK_2022_paper_5584}
    {0cm}
    {}
\addpaper
    {this is not a review}
    {Huilian Sophie Qiu}
    {Keywords: this, is, not, review}
    {SIGBOVIK_2022_paper_9376}
    {0cm}
    {}