### 2022-04-08 message-from-committee.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\usepackage{tipa}
\usepackage[normalem]{ulem}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\newcommand{\auth}{\text{auth}}
\newcommand{\paperby}{\text{paperBy}}
\newcommand{\reviewofby}{\text{reviewOfBy}}
\newcommand{\bool}{\text{bool}}
\newcommand{\rev}{\text{rev}}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2022}} \\ &\\
& \LARGE{Message from the Organizing Committee} \\
&\\
\hline
\end{tabular}}
\vspace{2em}
\thispagestyle{empty}

%For generality’s sake, we have templatized the message of the organizing committee. The actual message may be produced by running the TeX command at the end.
%
%\begin{verbatim}
%\newcommand{\Message2020}[3]{
%% TODO: generalize ordinal indicators
%\end{verbatim}

%Whoops! Looks like you found the placeholder message from the committee :( Don't worry, I'm sure someone next year will use this as the set up for a very funny joke.
%\begin{flushright}
%The SIGBOVIK 2021 Organizing Committee\\
%Pittsburgh, PA \\
%\& Online from Several Locations
%\vspace{1em}

%\begin{tabular}{r r p{0.5\linewidth}}
%	Asher Trockman (general chair) &
%Jenny Lin (easy chair)\\
%	Siva Somayyajula (senior hard-ass chair) &
%Sol Boucher (acting emeritus proceedings chair)\\
%	Brandon Bohrer (beanbag chair) &
%Ryan Kavanagh (rockin' chair)\\
%	Stefan Muller (ergonomic office chair) &
%Chris Yu (art chair)\\
%	Hana Frluckaj (moderation chair) &
%Daniel Smullen (moderation chair)\\
%	Xindi Wu (conference chair) &
%Sydney Gibson (tweet chair)\\
%	John Grosen (archaeology chair) &
%	Vivian Shen (honorary awards chair)
%\end{tabular}

%\end{flushright}
\begin{verbatim}

\Message2020{16}{uenchiest}{$2^6$}
\end{verbatim}

%\bibliographystyle{acm}
%\bibliography{msgrefs}
\thispagestyle{empty}

\end{document}