### 2023-04-05 papers.tex

\renewcommand{\thepage}{\roman{page}}
\setcounter{page}{3}

\newpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\addproceedings{THE ULTIMATE THROWBACK}

\newcommand{\link}[2][]{\vspace{-1in} {\scriptsize #1 \url{#2}} \raisebox{-0.5\height}{\includegraphics[height=0.75in]{participation}}}
\newcommand{\badge}[1]{\includegraphics[height=0.75in]{badges/badge_#1}}
\newcommand{\badges}[1]{\vspace{-1in} \raisebox{-0.5\height}{#1}}

\addtrack{A}{Multiplicity, Meet Singularity}
\addpaper
	{An Undergrad Is All You Need}
	{James Yoo}
	{}
	{SIGBOVIK_2023_paper_6.pdf}
	{0cm}
	{}
\addpaper
	{Transformers are robots in disguise but also:}
	{\protect{Michael Saxon, Luca Soldaini, Alexander F Kratz, ``, ''.join([f``\\textbf\{x\}'' for x in ALL\_SHUTAI\_EMPLOYEES])\}, *Optimus Prime [0x7c0], and David S. Hippocampus [0x1b39]}}
	{}
	{SIGBOVIK_2023_paper_15.pdf}
	{0cm}
	{}
\addpaper
	{The Implications of Sentient Chatbots}
	{Clark Levi Jones}
	{}
	{SIGBOVIK_2023_paper_9.pdf}
	{0cm}
	{}
\addpaper
	{AyahuascaNet: Rigorously Investigating Hallucination in Large Language Models with Hardcore Psychedelic Drugs}
	{Andre Ye}
	{}
	{SIGBOVIK_2023_paper_26.pdf}
	{0cm}
	{}
\addpaper
	{SocietyZoo: Exploring Anthropomorphic Traits in Diverse and Ingenious Neural Network Architectures}
	{Tarun Raheja and Nilay Pochhi}
	{}
	{SIGBOVIK_2023_paper_29.pdf}
	{0cm}
	{}
\addpaper
	{GradIEEEnt half decent}
	{Dr. Tom Murphy VII Ph.D.}
	{}
	{SIGBOVIK_2023_paper_32.pdf}
	{0cm}
	{}
\addpaper
	{Leveraging insect populations to implement large scale deep learning}
	{Aditi Kabra and Sagar Bharadwaj}
	{}
	{SIGBOVIK_2023_paper_40.pdf}
	{0cm}
	{}
\addpaper
	{Quantifying and Predicting Large Language Model Hype in SIGBOVIK and Beyond}
	{Wang, Kevin A., Khosravi, Pasha, Khosravi, Pooya, Chu, Linh, and Gajulapalli, Karthik}
	{}
	{SIGBOVIK_2023_paper_51.pdf}
	{0cm}
	{}
\addpaper
	{Unstable Diffusion: A Generative Model That Does Not Really Care}
	{Woddis Updog}
	{}
	{SIGBOVIK_2023_paper_55.pdf}
	{0cm}
	{}
\addpaper
	{You Won't Believe This One WEIRD TRICK That BEATS ChatGPT on AI{\small c} (NOT CLICKBAIT)}
	{Alex Xie, Abhishek Vijayakumar, Erin Gao, Bhargav Hadya, Samiksha Kale, and Tara Lakdawala}
	{}
	{SIGBOVIK_2023_paper_61.pdf}
	{0cm}
	{}
\addpaper
	{Alcatrez: A Large Language Model to Jailbreak Large Language Models}
	{}
	{}
	{SIGBOVIK_2023_paper_66.pdf}
	{0cm}
	{}
\addpaper
	{Meat-Based Graphics Pipelines}
	{Will BL}
	{}
	{SIGBOVIK_2023_paper_33.pdf}
	{0cm}
	{}

\addtrack{B}{Well-known Problem, Meet Solution}
\addpaper
	{Airport Security, Generalized Chess, and NP /= P}
	{Albert Gao}
	{}
	{SIGBOVIK_2023_paper_2.pdf}
	{0cm}
	{}
\addpaper
	{A Jalgorithm for Japplying Jeans to Jobjects}
	{Dave Pagurek}
	{}
	{SIGBOVIK_2023_paper_4.pdf}
	{0cm}
	{}
\addpaper
	{On the Origin of Sandwiches: A Revised Theory}
	{Dr. Esquire ZeWaka and Dr. ``The Big'' Jablonski}
	{}
	{SIGBOVIK_2023_paper_21.pdf}
	{0cm}
	{}
\addpaper
	{Is the number of Falco lasers ever shot greater than the number of humans alive?}
	{Arya Amami}
	{}
	{SIGBOVIK_2023_paper_36.pdf}
	{0cm}
	{}
\addpaper
	{Maximizing Code Readability Using Semicolon Indentation}
	{Tessa Parker and Nelitha Kulasiri}
	{}
	{SIGBOVIK_2023_paper_64.pdf}
	{0cm}
	{}
\addpaper
	{A perpetual motion machine}
	{Zikuan Wang}
	{}
	{SIGBOVIK_2023_paper_50.pdf}
	{0cm}
	{}

\addtrack{C}{Church, Meet State}
\addpaper
	{Even Lower Order Functions for Returning}
	{Saul Field, Dann Toliver, and Erik Derohanian }
	{}
	{SIGBOVIK_2023_paper_52.pdf}
	{0cm}
	{}
\addpaper
	{PizzaLang: A Language for Hungry Developers}
	{\protect{Ronik Bhaskar [0x7e7] and Dawson Ren [0x7e7]}}
	{}
	{SIGBOVIK_2023_paper_8.pdf}
	{0cm}
	{}
\addpaper
	{Quadratic Logic}
	{Daniel Ng}
	{}
	{SIGBOVIK_2023_paper_18.pdf}
	{0cm}
	{}

\addpaper
	{Ringfuck: It's a wheel great time!}
	{Ryan Hornby and Matthew Safar}
	{}
	{SIGBOVIK_2023_paper_39.pdf}
	{0cm}
	{}
\addpaper
	{NovelML: A Novel SML Semantics}
	{\protect{Thea Brick [0x00007e7] and James *Gallicchio [0x07e7]}}
	{}
	{SIGBOVIK_2023_paper_54.pdf}
	{0cm}
	{}
\addpaper
	{A Halt-Averse Instruction Set Architecture for Embedded Hypercomputers}
	{\protect{Ben Burrill [0x7F6, 0x76D]}}
	{}
	{SIGBOVIK_2023_paper_27.pdf}
	{0cm}
	{}

\addtrack{D}{Complexity, Meet Simplicity}
\addpaper
	{Simultaneous Paper Maximization and Minimization Through Reference List Side Channel Information Injection}
	{Frans Skarman}
	{}
	{SIGBOVIK_2023_paper_43.pdf}
	{0cm}
	{}
\addpaper
	{On the Turing Completeness of Eeny, Meeny, Miny, Moe}
	{Javier Lim}
	{}
	{SIGBOVIK_2023_paper_44.pdf}
	{0cm}
	{}

\addtrack{E}{Publishing, Meet Perishing}
\addpaper
	{TeX-to-TikZ}
	{Alex Friedman, Charlotte Clark, Natalie McClain, Shaw-leon Chen, Connie Chau, Jake Backer, N'yoma Diamond, and Rose Bohrer}
	{}
	{SIGBOVIK_2023_paper_13.pdf}
	{0cm}
	{}
\addpaper
	{Cerberus: Why Papers Named Cerberus Always Get Accepted}
	{Ben Weintraub}
	{}
	{SIGBOVIK_2023_paper_28.pdf}
	{0cm}
	{}
\addpaper
	{Large Language Models are Few-Shot Publication Scoopers}
	{Samuel Albanie, Liliane Momeni, and Joao F. Henriques}
	{}
	{SIGBOVIK_2023_paper_42.pdf}
	{0cm}
	{}
\addpaper
	{Author-Unification: Name-, Institution-, and Career-Sharing Co-authors}
	{Vanessa Wirth, Vanessa Wirth}
	{}
	{SIGBOVIK_2023_paper_45.pdf}
	{0cm}
	{}

\addtrack{F}{Fun, Meet Games}
\addpaper
	{Multidiscipline Elo - A Complex Analysis}
	{Kiera Jones}
	{}
	{SIGBOVIK_2023_paper_20.pdf}
	{0cm}
	{}
\addpaper
	{A Simple RISK Architecture}
	{Stefan Muller}
	{}
	{SIGBOVIK_2023_paper_31.pdf}
	{0cm}
	{}
\addpaper
	{Solving Catan}
	{Sam Triest, Sam Schoedel}
	{}
	{SIGBOVIK_2023_paper_56.pdf}
	{0cm}
	{}
\addpaper
	{Code Golfing for Characters (not Bytes)}
	{Russell Schwartz}
	{}
	{SIGBOVIK_2023_paper_10.pdf}
	{0cm}
	{}

\addtrack{G}{Sharing, Meet Caring}
\addpaper
	{Screen-sharing Concurrency}
	{Screen-Sharing Concurrency}
	{}
	{SIGBOVIK_2023_paper_11.pdf}
	{0cm}
	{}
\addpaper
	{Fair Division of a Single Indivisible Object}
	{Gabriel Chuang}
	{}
	{SIGBOVIK_2023_paper_12.pdf}
	{0cm}
	{}
\addpaper
	{Stretch Goals: Gamification of SLINKYs}
	{Geoff Barnes, William Gunther, and Brian Kell}
	{}
	{SIGBOVIK_2023_paper_59.pdf}
	{0cm}
	{}

\addtrack{H}{Post-Quantum, Meet Ergo Propter-Quantum}
\addpaper
	{Voynichese Cryptography}
	{Roger Bacon}
	{}
	{SIGBOVIK_2023_paper_19.pdf}
	{0cm}
	{}
\addpaper
	{Coupon Code Generation: Saving space with a simple (and insecure) hashing technique}
	{Ben Rosenberg}
	{}
	{SIGBOVIK_2023_paper_3.pdf}
	{0cm}
	{}
\addpaper
	{Natural Differential Privacy}
	{Micah Altman and Aloni Cohen}
	{}
	{SIGBOVIK_2023_paper_46.pdf}
	{0cm}
	{}
\addpaper
	{From Zero to Hero: Convincing with Extremely Complicated Math}
	{Maximilian Weiherer and Bernhard Egger}
	{}
	{SIGBOVIK_2023_paper_47.pdf}
	{0cm}
	{}
\addpaper
	{Quantum Bogosort}
	{George Huebner and Beam Fujimoto}
	{}
	{SIGBOVIK_2023_paper_49.pdf}
	{0cm}
	{}

\addtrack{I}{Line of Inquiry, Meet Your Logical Conclusion}
\addpaper
	{Unlimited null: achieving memory safety by extending memory protection}
	{hikari\_no\_yume}
	{}
	{SIGBOVIK_2023_paper_14.pdf}
	{0cm}
	{}
\addpaper
	{miles2km: The worst ways to convert from miles to km}
	{Lucien Rae}
	{}
	{SIGBOVIK_2023_paper_30.pdf}
	{0cm}
	{}
\addpaper
	{Alan Hsu}
	{Alex Xie and Alan Hsu}
	{}
	{SIGBOVIK_2023_paper_35.pdf}
	{0cm}
	{}
\addpaper
	{ACHOO: Actually Higher Order Optimization}
	{Clayton W Thorrez}
	{}
	{SIGBOVIK_2023_paper_57.pdf}
	{0cm}
	{}
\addpaper
	{The Phonetic Portmantout}
	{Kyle A. Williams}
	{}
	{SIGBOVIK_2023_paper_58.pdf}
	{0cm}
	{}
\addpaper
	{An Introduction to Compliers}
	{Anoushka Shrivastava}
	{}
	{SIGBOVIK_2023_paper_63.pdf}
	{0cm}
	{}

\addpaper
	{Feline Fine: A Purr-spicacious Proof of Nekomusume Supremacy Over Human Females}
	{\protect{Dr. Tuesday [03745] and Hunter Halloran}}
	{}
	{SIGBOVIK_2023_paper_24.pdf}
	{0cm}
	{}

\addtrack{J}{New York Times, Meet Your Next Cash Cow}
\addpaper
	{Elo Worldle, a framework for teaching the children about weak chess engines}
	{Clark Levi Jones}
	{}
	{SIGBOVIK_2023_paper_37.pdf}
	{0cm}
	{}
\addpaper
	{Harderdl: Yet another Wordle variation for those who like challenges}
	{Elkin Cruz-Camacho}
	{}
	{SIGBOVIK_2023_paper_60.pdf}
	{0cm}
	{}

\addtrack{K}{SIGBOVIK, Meet Your Match}
\addpaper
	{Bulletin Board Sigbovik}
	{Tianjian Hu}
	{}
	{SIGBOVIK_2023_paper_17.pdf}
	{0cm}
	{}
\addpaper
	{Poisoning SIGBOVIK-Scale Training Datasets is Practical}
	{Tobias Pfandzelter}
	{}
	{SIGBOVIK_2023_paper_22.pdf}
	{0cm}
	{}
\addpaper
	{\emph{($\star$ AGI Track):} Poisoning SIGBOVIK-Scale Training Datasets is Practical}
	{Tobias Pfandzelter}
	{}
	{SIGBOVIK_2023_paper_23.pdf}
	{0cm}
	{}
\addpaper
	{Rizz Is All You Need: Expert Dating via Reinforcement Learning}
	{Fan Pu Zeng}
	{}
	{SIGBOVIK_2023_paper_65.pdf}
	{0cm}
	{}

\addtrack{L}{Hunger, Meet Pickle}
\addpaper
	{The Influence of Lunch Items on Cryptocurrency in the United States}
	{\protect {*David An [0x7e7]}}
	{}
	{SIGBOVIK_2023_paper_7.pdf}
	{0cm}
	{}
\addpaper
	{}
	{}
	{}
	{SIGBOVIK_2023_paper_41.pdf}
	{0cm}
	{}
\addpaper
	{Tactical Toast Cut Silhouette Recognition Guide}
	{Jim McCann}
	{}
	{SIGBOVIK_2023_paper_62.pdf}
	{0cm}
	{}
\addpaper
	{Saltgurken: A formal grammar for unambiguous grocery shopping}
	{*Eleftheria Chatziargyriou}
	{}
	{SIGBOVIK_2023_paper_34.pdf}
	{0cm}
	{}

\addtrack{M}{Hear, Meet And Now}
\addpaper
	{VOACaloid: A ``better'' ``hardware-based'' ``portable'' ``solution'' for the ``real-time'' ``generation'' of ``singing''}
	{Aaron Thapa}
	{}
	{SIGBOVIK_2023_paper_38.pdf}
	{0cm}
	{}
\addpaper
	{Avantgarde Visual Auditive JSON Hashing}
	{Tobias Coq, Louis Erhard, Julio Esperantò, Jean Parler, and Peter-Ludwig Sechsdiener}
	{}
	{SIGBOVIK_2023_paper_48.pdf}
	{0cm}
	{}

\addtrack{N}{Reader, Meet Remainder}
%withdrawn:
%\addpaper
%	{Estimating the Probability That the Explosion of an Ink Sphere Produces a Dictionary}
%	{Braden Oh}
%	{}
%	{papers/SIGBOVIK_2023_paper_1.pdf}
%	{0cm}
%	{}
\addpaper
	{New Advancements in How Fucked You Are if You Don’t Use Our Software}
	{\protect{Lemma Brummel and Society for Creating Amplified Monetizability in Augmented Industries [0x202b]}}
	{}
	{SIGBOVIK_2023_paper_25.pdf}
	{0cm}
	{}
\addpaper
	{A Retrospective Psychological Evaluation of the Logical Contradictions in Writing Systems Containing Japanese Kanji and Chinese Characters}
	{Ashvin Ranjan}
	{}
	{SIGBOVIK_2023_paper_53.pdf}
	{0cm}
	{}
\addpaper
	{Health Code: COVID Control and Advancements in Digital Image Compression}
	{Tianjian Hu}
	{}
	{SIGBOVIK_2023_paper_5.pdf}
	{0cm}
	{}