### 2023-04-05 titlepage.tex

% NB. xelatex this file.

\documentclass{article}
\usepackage{fontspec}

\mathchardef\mhyphen="2D

\begin{document}

\setmainfont{Splotch.ttf}
\setsansfont{Birbaslo.ttf}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\begin{titlepage}
\begin{center}

% Upper part of the page. The '~' is needed because \\
% only works if a paragraph has started.
\textsc{\LARGE the association for computational heresy}\\[0.5cm]

\textsc{\large presents}\\[1cm]

\textsc{\Large a record of the proceedings of}\\[0.5cm]
% Title
\HRule \\[0.4cm]
% TODO: Don't listen to copyright-page.tex: this also needs to be updated every year!
{ \Huge \sf SIGBOVIK 0x2023 \\[0.4cm] }

{\fontspec{Birbaslo.ttf} THE ULTIMATE THROWBACK}

\HRule \\[2cm]

% Author and supervisor

% TODO: So does this!
the {\fontspec{Birbaslo.ttf} \em last} annual intercalary robot dance party in celebration

of workshop on symposium about $2^6$th birthdays{\fontspec{Birbaslo.ttf} ;} in particular{\fontspec{Birbaslo.ttf} ,}

that of harry q{\fontspec{Birbaslo.ttf} .} bovik

\vfill

cover art by entity rosebohrercs
% TODO: Remove this, if applicable.
%global chaos courtesy of sars$\mhyphen$cov$\mhyphen 2$

\vfill

% Bottom of the page
{\Large carnegie mellon university\\[0.25cm]
pittsburgh{\fontspec{Birbaslo.ttf} ,} pa\\[0.3cm]
% TODO: And this one!
april $0${\fontspec{Birbaslo.ttf} ,} $2023$}

\end{center}
\end{titlepage}

\end{document}