### 2024-04-07 message-from-committee.tex
2024 init commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\usepackage{tipa}
\usepackage[normalem]{ulem}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\newcommand{\auth}{\text{auth}}
\newcommand{\paperby}{\text{paperBy}}
\newcommand{\reviewofby}{\text{reviewOfBy}}
\newcommand{\bool}{\text{bool}}
\newcommand{\rev}{\text{rev}}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 0x2024}} \\ &\\
& \LARGE{Message from the Organizing Committee} \\
&\\
\hline
\end{tabular}}
\vfill
\thispagestyle{empty}

%For generality’s sake, we have templatized the message of the organizing committee. The actual message may be produced by running the TeX command at the end.
%
%\begin{verbatim}
%\newcommand{\Message2020}[3]{
%% TODO: generalize ordinal indicators
%\end{verbatim}

%Whoops! Looks like you found the placeholder message from the committee :( Don't worry, I'm sure someone next year will use this as the set up for a very funny joke.
%\begin{flushright}
%The SIGBOVIK 2021 Organizing Committee\\
%Pittsburgh, PA \\
%\& Online from Several Locations
%\vspace{1em}

%\begin{tabular}{r r p{0.5\linewidth}}
%	Asher Trockman (general chair) &
%Jenny Lin (easy chair)\\
%	Siva Somayyajula (senior hard-ass chair) &
%Sol Boucher (acting emeritus proceedings chair)\\
%	Brandon Bohrer (beanbag chair) &
%Ryan Kavanagh (rockin' chair)\\
%	Stefan Muller (ergonomic office chair) &
%Chris Yu (art chair)\\
%	Hana Frluckaj (moderation chair) &
%Daniel Smullen (moderation chair)\\
%	Xindi Wu (conference chair) &
%Sydney Gibson (tweet chair)\\
%	John Grosen (archaeology chair) &
%	Vivian Shen (honorary awards chair)
%\end{tabular}

%\end{flushright}
%\begin{verbatim}
%\Message2020{16}{uenchiest}{$2^6$}
%\end{verbatim}

time is small.

brevity is the wit of soul.

SIGBOVIK 2024 was full of plusplusgood articles.

A.C., H.P., J.M., T.7, S.B., S.M., and H.Q.B. were actively involved in publication and organization.

thus marks the end of the proceedings for this year.

-the SIGBOVIK '24/7\\

\vfill

%\bibliographystyle{acm}
%\bibliography{msgrefs}
\thispagestyle{empty}

\end{document}